---
title: "Loperamide 16S V6 amplicons analysis"
author: "Rebecca Stevick"
date: "7/1/2022"
output:
  html_document:
    toc: true
    keep_md: TRUE
    theme: "cerulean"
    toc_float:
      collapsed: false
      smooth_scroll: false
---

# About the Data

## Conditions

During Trial 38.

Conventional fish treated with loperamide at 5 dpf for 24 hours.

-   T0. Sample timepoint 1 at 6 dpf (24 hour treatment)
-   T1. Sample timepoint 2 at 7 dpf (24 hour treatment + 24 hour water)
-   T5. Sample timepoint 3 at 11 dpf (24 hour treatment + 5 days water)


| ID |   SeqName   | SampleName |   Project  | FishType | LoperamidePathogen | SampleRep | Timepoint | TimepointDay |
|:--:|:-----------:|:----------:|:----------:|:--------:|:------------------:|:---------:|:---------:|:------------:|
| 1  | 1_Conv_1_1  | Conv_1_1   | Loperamide | Conv     |                    | 1         | 1         | 24h          |
| 2  | 2_Conv_1_2  | Conv_1_2   | Loperamide | Conv     |                    | 2         | 1         | 24h          |
| 3  | 3_Conv_1_3  | Conv_1_3   | Loperamide | Conv     |                    | 3         | 1         | 24h          |
| 4  | 4_Conv_1_4  | Conv_1_4   | Loperamide | Conv     |                    | 4         | 1         | 24h          |
| 5  | 5_Conv_1_5  | Conv_1_5   | Loperamide | Conv     |                    | 5         | 1         | 24h          |
| 6  | 6_DMSO_1_1  | DMSO_1_1   | Loperamide | Conv     | DMSO 1:100         | 1         | 1         | 24h          |
| 7  | 7_DMSO_1_2  | DMSO_1_2   | Loperamide | Conv     | DMSO 1:100         | 2         | 1         | 24h          |
| 8  | 8_DMSO_1_3  | DMSO_1_3   | Loperamide | Conv     | DMSO 1:100         | 3         | 1         | 24h          |
| 9  | 9_DMSO_1_4  | DMSO_1_4   | Loperamide | Conv     | DMSO 1:100         | 4         | 1         | 24h          |
| 10 | 10_DMSO_1_5 | DMSO_1_5   | Loperamide | Conv     | DMSO 1:100         | 5         | 1         | 24h          |
| 11 | 11_LOP_1_1  | LOP_1_1    | Loperamide | Conv     | Loperamide 10mg/L  | 1         | 1         | 24h          |
| 12 | 12_LOP_1_2  | LOP_1_2    | Loperamide | Conv     | Loperamide 10mg/L  | 2         | 1         | 24h          |
| 13 | 13_LOP_1_3  | LOP_1_3    | Loperamide | Conv     | Loperamide 10mg/L  | 3         | 1         | 24h          |
| 14 | 14_LOP_1_4  | LOP_1_4    | Loperamide | Conv     | Loperamide 10mg/L  | 4         | 1         | 24h          |
| 15 | 15_LOP_1_5  | LOP_1_5    | Loperamide | Conv     | Loperamide 10mg/L  | 5         | 1         | 24h          |
| 16 | 16_Conv_2_1 | Conv_2_1   | Loperamide | Conv     |                    | 1         | 2         | 48h          |
| 17 | 17_Conv_2_2 | Conv_2_2   | Loperamide | Conv     |                    | 2         | 2         | 48h          |
| 18 | 18_Conv_2_3 | Conv_2_3   | Loperamide | Conv     |                    | 3         | 2         | 48h          |
| 19 | 19_Conv_2_4 | Conv_2_4   | Loperamide | Conv     |                    | 4         | 2         | 48h          |
| 20 | 20_Conv_2_5 | Conv_2_5   | Loperamide | Conv     |                    | 5         | 2         | 48h          |
| 21 | 21_DMSO_2_1 | DMSO_2_1   | Loperamide | Conv     | DMSO 1:100         | 1         | 2         | 48h          |
| 22 | 22_DMSO_2_2 | DMSO_2_2   | Loperamide | Conv     | DMSO 1:100         | 2         | 2         | 48h          |
| 23 | 23_DMSO_2_3 | DMSO_2_3   | Loperamide | Conv     | DMSO 1:100         | 3         | 2         | 48h          |
| 24 | 24_DMSO_2_4 | DMSO_2_4   | Loperamide | Conv     | DMSO 1:100         | 4         | 2         | 48h          |
| 25 | 25_DMSO_2_5 | DMSO_2_5   | Loperamide | Conv     | DMSO 1:100         | 5         | 2         | 48h          |
| 26 | 26_LOP_2_1  | LOP_2_1    | Loperamide | Conv     | Loperamide 10mg/L  | 1         | 2         | 48h          |
| 27 | 27_LOP_2_2  | LOP_2_2    | Loperamide | Conv     | Loperamide 10mg/L  | 2         | 2         | 48h          |
| 28 | 28_LOP_2_3  | LOP_2_3    | Loperamide | Conv     | Loperamide 10mg/L  | 3         | 2         | 48h          |
| 29 | 29_LOP_2_4  | LOP_2_4    | Loperamide | Conv     | Loperamide 10mg/L  | 4         | 2         | 48h          |
| 30 | 30_LOP_2_5  | LOP_2_5    | Loperamide | Conv     | Loperamide 10mg/L  | 5         | 2         | 48h          |
| 31 | 31_Conv_3_1 | Conv_3_1   | Loperamide | Conv     |                    | 1         | 3         | 6d           |
| 32 | 32_Conv_3_2 | Conv_3_2   | Loperamide | Conv     |                    | 2         | 3         | 6d           |
| 33 | 33_Conv_3_3 | Conv_3_3   | Loperamide | Conv     |                    | 3         | 3         | 6d           |
| 34 | 34_Conv_3_4 | Conv_3_4   | Loperamide | Conv     |                    | 4         | 3         | 6d           |
| 35 | 35_Conv_3_5 | Conv_3_5   | Loperamide | Conv     |                    | 5         | 3         | 6d           |
| 36 | 36_DMSO_3_1 | DMSO_3_1   | Loperamide | Conv     | DMSO 1:100         | 1         | 3         | 6d           |
| 37 | 37_DMSO_3_2 | DMSO_3_2   | Loperamide | Conv     | DMSO 1:100         | 2         | 3         | 6d           |
| 38 | 38_DMSO_3_3 | DMSO_3_3   | Loperamide | Conv     | DMSO 1:100         | 3         | 3         | 6d           |
| 39 | 39_DMSO_3_4 | DMSO_3_4   | Loperamide | Conv     | DMSO 1:100         | 4         | 3         | 6d           |
| 40 | 40_DMSO_3_5 | DMSO_3_5   | Loperamide | Conv     | DMSO 1:100         | 5         | 3         | 6d           |
| 41 | 41_LOP_3_1  | LOP_3_1    | Loperamide | Conv     | Loperamide 10mg/L  | 1         | 3         | 6d           |
| 42 | 42_LOP_3_2  | LOP_3_2    | Loperamide | Conv     | Loperamide 10mg/L  | 2         | 3         | 6d           |
| 43 | 43_LOP_3_3  | LOP_3_3    | Loperamide | Conv     | Loperamide 10mg/L  | 3         | 3         | 6d           |
| 44 | 44_LOP_3_4  | LOP_3_4    | Loperamide | Conv     | Loperamide 10mg/L  | 4         | 3         | 6d           |
| 45 | 45_LOP_3_5  | LOP_3_5    | Loperamide | Conv     | Loperamide 10mg/L  | 5         | 3         | 6d           |



## Sampling

1.  Transfer fish into a small petri dish full of sterile water (wash \#1).
2.  Wash the fish again by transferring into a new petri dish full of sterile water (wash \#2).
3.  Add 5 washed fish per tube into 5 x 2 mL tubes, with \~500 µL of water
4.  After all samples have been collected for all conditions, remove as much water as possible with a p200 pipette
5.  Put the samples at -80C

## Sequencing

16S V6 region

*Primers*  

1064R GTCTCGTGGGCTCGGAGATGTGTATAAGAGACAGCGACRRCCATGCANCACCT   
967F-A TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGCTAACCGANGAACCTYACC   
967F-B TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGCNACGCGAAGAACCTTANC    
967F-C TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGCAACGCGMARAACCTTACC    
967F-D TCGTCGGCAGCGTCAGATGTGTATAAGAGACAGATACGCGARGAACCTTACC

2 x 250 bp paired-end Illumina sequencing

## Analysis

Initially done with QIIME2, dada2 denoising. Eukaryotic reads removed here.

# Setup

## Load libraries and settings

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(scales)
library(ggpubr)
#library(ungeviz)
library(qiime2R)
library(vegan)
library(gt)
library(microbiomeMarker)
library(patchwork)
library(ggh4x)
library(ggtext)

# set global theme
theme_set(theme_minimal()+
             theme(panel.grid.major.y = element_line(color="grey80"), strip.text=element_text(size=16),
                   strip.background = element_rect(color="transparent"),
                   strip.text.y = element_text(angle=0), plot.caption = element_text(size=10),
                   panel.grid.major.x = element_blank(),legend.position="top",
                   plot.background = element_rect(fill="transparent", color="transparent"),
                   axis.ticks = element_line(inherit.blank = FALSE),
                   panel.background = element_rect(color="grey50", size=2), 
                   legend.title = element_text(size=18),
                   axis.text = element_text(size=15), axis.title = element_text(size=18),
                   legend.text = element_text(size=16), plot.title = element_text(hjust=0.5)))

# set global options for code output
knitr::opts_chunk$set(echo=TRUE, warning=FALSE,message=FALSE)

```

## Import data

```{r import}

metadatafile <- "YersiniaLoperamide16Samplicon_metadata.txt"
tablefile <- "table.qza"
taxonomyfile <- "taxonomy.qza"
treefile <- "rooted-tree.qza"

SVs<-read_qza(tablefile)$data
head(SVs)[,1:5]

metadata<-read_q2metadata(metadatafile)  %>% 
   mutate(Loperamide=case_when(LoperamidePathogen == "DMSO 1:100" ~ "DMSO",
                                       LoperamidePathogen == "Loperamide 10mg/L" ~ "Loperamide",
                                       TRUE ~ "Control water"))
head(metadata)[,1:5]

taxonomy<-read_qza(taxonomyfile)$data %>% parse_taxonomy()
head(taxonomy)[,1:5]

physeq<-qza_to_phyloseq(
    features=tablefile,
    tree=treefile,
    taxonomy=taxonomyfile,
    metadata = metadatafile
    )

```

## Clean up data

```{r summarize}

metadata$LoperamideTimepoint <- factor(metadata$LoperamideTimepoint, 
                          levels = c("ControlWater.24h","ControlWater.48h","ControlWater.6d",
                                     "DMSO.24h", "DMSO.48h","DMSO.6d",
                                     "Loperamide.24h","Loperamide.48h","Loperamide.6d"),
                          labels = c("ControlWater.T0","ControlWater.T1","ControlWater.T5",
                                     "DMSO.T0", "DMSO.T1","DMSO.T5",
                                     "Loperamide.T0","Loperamide.T1","Loperamide.T5"))

metadata$TimepointDay <- factor(metadata$TimepointDay, 
                          levels = c("24h","48h","6d"),
                          labels = c("T0","T1","T5"))


metadataLoperamide <- filter(metadata, Project=="Loperamide" | Project=="Controls")
taxasums <- qiime2R::summarize_taxa(SVs, taxonomy) 

# vector of eukaryotic & cyanobacteria ASVs
eukASVs <- taxonomy %>% filter(Kingdom=="d__Eukaryota" | Kingdom=="d__Archaea" | Phylum=="Cyanobacteria") %>% 
   rownames_to_column("ASVs") %>% .$ASVs
# just taxa data in loepramide study, without archea/chloroplasts/eukaryotic
taxasumsclean <- select(as.data.frame(SVs), metadataLoperamide$SampleID) %>% 
   rownames_to_column("ASV") %>% 
   filter(!ASV %in% eukASVs) %>% 
   column_to_rownames(var="ASV") %>% qiime2R::summarize_taxa(taxonomy) 

datafullphylum <- 
   # transpose the phylum table of taxa sums
   t(taxasumsclean$Phylum) %>% 
   # make into data frame
   as.data.frame() %>% 
   # put rownames as SampleID
   rownames_to_column("SampleID") %>% 
   # join with the metadata
   full_join(metadata) %>% 
   # convert to long format
   pivot_longer("d__Bacteria; Abditibacteriota":"Unassigned; NA")

datafullkingdom <- 
   t(taxasumsclean$Kingdom) %>%  as.data.frame() %>% rownames_to_column("SampleID") %>% 
   full_join(metadata) %>% pivot_longer("d__Bacteria":"Unassigned")

datafullclass <- 
   t(taxasumsclean$Class) %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
   full_join(metadata) %>% 
   pivot_longer("d__Bacteria; Abditibacteriota; Abditibacteria":"Unassigned; NA; NA")

datafullorder <- 
   t(taxasumsclean$Order) %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
   full_join(metadata) %>% 
   pivot_longer("d__Bacteria; Abditibacteriota; Abditibacteria; Abditibacteriales":"Unassigned; NA; NA; NA")

datafullfamily <- 
   t(taxasumsclean$Family) %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
   full_join(metadata) %>% 
   pivot_longer("d__Bacteria; Abditibacteriota; Abditibacteria; Abditibacteriales; Abditibacteriaceae":
                   "Unassigned; NA; NA; NA; NA")

datafullgenus <- 
   t(taxasumsclean$Genus) %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
   full_join(metadata) %>% 
   pivot_longer("d__Bacteria; Abditibacteriota; Abditibacteria; Abditibacteriales; Abditibacteriaceae; Abditibacterium":
                   "Unassigned; NA; NA; NA; NA; NA")

datafullASVs <- t(SVs) %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
   full_join(metadata) %>% 
   pivot_longer("cca00a58137be2c40384b70e46bda58e":"213dc85b14c8cdc3e765c92bcc6fa6a2", names_to = "ASV") %>% 
   filter(Project=="Loperamide" | Project=="Controls") %>% 
   filter(!ASV %in% eukASVs)


# matrix format
metadataLoperamide <- filter(metadata, Project=="Loperamide")
matrixLoperamide <- select(as.data.frame(SVs), metadataLoperamide$SampleID) %>% rownames_to_column("ASV") %>% 
   filter(!ASV %in% eukASVs) %>% 
   column_to_rownames(var="ASV") %>% t()

matrixLoperamidePercent <- select(as.data.frame(SVs), metadataLoperamide$SampleID) %>% rownames_to_column("ASV") %>% 
   filter(!ASV %in% eukASVs) %>% 
   column_to_rownames(var="ASV") %>% 
   mutate_if(is.numeric, proportions) %>% t()

# physeq format
Loperamidephyseq <- phyloseq::subset_samples(physeq, Project=="Loperamide") %>% 
 #  phyloseq::subset_samples(LoperamidePathogen=="DMSO 1:100" | LoperamidePathogen=="Loperamide 10mg/L") %>% 
   phyloseq::subset_taxa((Kingdom!="d__Eukaryota") | is.na(Kingdom)) %>% 
   phyloseq::subset_taxa((Kingdom!="d__Archaea") | is.na(Kingdom)) %>% 
   phyloseq::subset_taxa((Phylum!="Cyanobacteria") | is.na(Phylum))

length(Loperamidephyseq@tax_table)


# datafullgenusf <- t(taxasums$Genus) %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
#    full_join(metadata) %>% 
#    filter(Project=="Loperamide" | Project=="Controls")
# 
#write.table(datafullgenusf, "datafullgenus.tsv", col.names = NA)
#write.table(datafullphylum, "datafullphylum.tsv")
#write.table(datafullfamily, "datafullfamily.tsv")



```


## Eukaryotic reads per sample



```{r eukreads, fig.width=16, fig.height=4}

t(taxasums$Phylum) %>% 
   as.data.frame() %>% 
   rownames_to_column("SampleID") %>% 
   full_join(metadata) %>% 
   pivot_longer("d__Archaea; Crenarchaeota":"Unassigned; NA") %>% 
   filter(Project=="Loperamide" | Project=="Controls") %>% 
   group_by(SampleName,Loperamide) %>% 
   mutate(percent=value/sum(value)) %>% 
   filter(name=="d__Eukaryota; Vertebrata") %>%
   ggplot(aes(x=SampleName, y=percent, fill=LoperamideTimepoint))+
   facet_nested(.~Project+Loperamide+TimepointDay, scales="free", space="free")+
   geom_col()+
   theme(legend.text = element_text(size=12, colour="gray20"),
         strip.background = element_rect(color="white", fill="grey85"),
         legend.position = "none",axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c("black","grey80","grey50","grey20","#a7a7ff",
                              "#0000ce", "#00006c","aquamarine1", "aquamarine3","aquamarine4"))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0), limits=c(0,1))+
   labs(y="Relative percent abundance",x=NULL,
        title="Relative percent abundance of d__Eukaryota; Vertebrata reads")

t(taxasums$Phylum) %>% 
   as.data.frame() %>% 
   rownames_to_column("SampleID") %>% 
   full_join(metadata) %>% 
   pivot_longer("d__Archaea; Crenarchaeota":"Unassigned; NA") %>% 
   filter(Project=="Loperamide" | Project=="Controls") %>% 
   filter(name=="d__Eukaryota; Vertebrata") %>%
   ggplot(aes(x=SampleName, y=value, fill=LoperamideTimepoint))+
   facet_nested(.~Project+Loperamide+TimepointDay, scales="free", space="free")+
   geom_col()+
   theme(legend.text = element_text(size=12, colour="gray20"),        
         strip.background = element_rect(color="white", fill="grey85"),
         legend.position = "none",axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c("black","grey80","grey50","grey20","#a7a7ff","#0000ce", 
                              "#00006c","aquamarine1", "aquamarine3","aquamarine4"))+
   scale_y_continuous(labels = scales::comma_format(), expand=c(0,0))+
   labs(y="Number of reads",x=NULL,
        title="Abundance of d__Eukaryota; Vertebrata reads")

```

-----


# General QC

## Reads per sample

```{r readnumberqc, fig.width=16, fig.height=4}

# define some distinct palettes
palette<-c("grey50","#D95F02", "#1B9E77", "#E6AB02","#7570B3")
palettep<-c("skyblue","#D95F02", "#1B9E77", "#E6AB02","#7570B3", "#66A61E", "gray30")
palettess<-c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F", "#FF7F00", "#CAB2D6", "#6A3D9A", "#FFFF99", "#B15928","gray58","grey40")

datafullkingdom %>% filter(Project=="Loperamide") %>% 
   group_by(SampleName, Project, Loperamide, TimepointDay) %>% summarise(NumReads = sum(value)) %>% 
   arrange(NumReads)

datafullkingdom %>% filter(Project=="Loperamide") %>% summarise(NumReads = sum(value))

readspersampleplot <- datafullkingdom %>% filter(Project=="Loperamide") %>% 
   ggplot(aes(x=factor(SampleName,levels=unique(metadata$SampleName)), 
              y=value, fill=Loperamide))+
   facet_nested(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_col()+
   theme(legend.position = "none", axis.text.x = element_text(size=10, angle=40, hjust=1), 
         strip.background = element_rect(color="white", fill="grey85"),
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
   scale_y_continuous(expand=c(0,0), labels=scales::label_comma())+
   labs(y="Number of reads per sample",x=NULL,fill=NULL)
readspersampleplot

```


## Rarefaction curves

```{r rarefaction}

(S <- specnumber(matrixLoperamide))
(raremax <- min(rowSums(matrixLoperamide)))
Srare <- vegan::rarefy(matrixLoperamide, raremax)
Sslope <- vegan::rareslope(matrixLoperamide, raremax)


plot(S, Srare, xlab = "Observed No. of ASVs", ylab = "Rarefied No. of ASVs")
abline(0,1)
plot(S, Sslope, xlab = "Observed No. of ASVs", ylab = "Slope at rarefied sample size")


rarecurve_data <- rarecurve(matrixLoperamide, step = 100, sample = raremax)

map_dfr(rarecurve_data, bind_rows) %>% 
   bind_cols(SampleID = rownames(matrixLoperamide),.) %>%
   pivot_longer(-SampleID) %>%
   drop_na() %>%
   mutate(n_seqs = as.numeric(str_replace(name, "N", ""))) %>%
   select(-name) %>%
   left_join(metadata) %>% 
   ggplot(aes(x=n_seqs, y=value, group=SampleID, color=LoperamideTimepoint, lty=Loperamide)) +
   geom_line(lwd=0.8) +
   scale_color_manual(values=c("grey80","grey50","grey20","#a7a7ff","#0000ce", 
                              "#00006c","aquamarine1", "aquamarine3","aquamarine4"))+
   scale_linetype_manual(values=c("solid","dotted","dashed"))+
   scale_x_continuous(labels=label_comma())+
   theme(legend.key.width = unit(1.33,"cm"), legend.position="right",
         strip.background = element_rect(color="white", fill="grey85"))+
   labs(x = "Number of sequences", y="Number of ASVs detected", color=NULL, lty=NULL)

rareplot <- map_dfr(rarecurve_data, bind_rows) %>% 
   bind_cols(SampleID = rownames(matrixLoperamide),.) %>%
   pivot_longer(-SampleID) %>%
   drop_na() %>%
   mutate(n_seqs = as.numeric(str_replace(name, "N", ""))) %>%
   select(-name) %>%
   left_join(metadata) %>% 
   ggplot(aes(x=n_seqs, y=value, group=SampleID, color=Loperamide, lty=Loperamide)) +
   geom_line(lwd=0.8) +
   facet_grid(.~TimepointDay)+
   scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
   scale_linetype_manual(values=c("solid","dotted","dashed"))+
   scale_x_continuous(labels=label_comma())+
   theme(legend.key.width = unit(1.33,"cm"), legend.position=c(0.9,0.18),
         strip.background = element_rect(color="white", fill="grey85"))+
#   geom_vline(xintercept = raremax, color="red")+
   labs(x = "Number of sequences", y="Number of ASVs detected", color=NULL, lty=NULL)

```

```{r qcfigure, fig.width=16, fig.height=10}

readspersampleplot / rareplot + plot_annotation(tag_levels = "A") &
   theme(plot.tag = element_text(face = "bold", size=20))

ggsave("LoperamideQC_rarecurve_reads.png", width=14, height=10)

```


## Controls

**Expected mock community**  


```{r mockcommunity, fig.width=8}

expectedmock <- data_frame(
   SampleName="Reality_Mock",
   value=0,
   TaxaOther=c("d3962c483c45f0765b9b7656ec7f43a0: \nd__Bacteria; p__Proteobacteria; c__Gammaproteobacteria",
               "fdaab5e761766ddd1297bd5f49d1cfb3: \nd__Bacteria; p__Proteobacteria; c__Gammaproteobacteria; o__Enterobacterales",
               "b08e8a180eba5a20a70b190fbf5f9ad2: \nd__Bacteria; p__Proteobacteria; c__Gammaproteobacteria; o__Enterobacterales; f__Enterobacteriaceae",
               "3019073bed0ceecaece9f446607c003a: \nd__Bacteria; p__Proteobacteria; c__Gammaproteobacteria; o__Enterobacterales; f__Enterobacteriaceae; g__Salmonella",
               "94efb59163996601f8a20a8f9b0ac0d9: \nd__Bacteria; p__Firmicutes; c__Bacilli; o__Bacillales; f__Bacillaceae; g__Bacillus",
               "456f366888eac2ce0809986d3ed1b237: \nd__Bacteria; p__Firmicutes; c__Bacilli; o__Lactobacillales",
               "d02f4c8904ba636c08bbdf8b65b5d7d0: \nd__Bacteria; p__Firmicutes; c__Bacilli; o__Lactobacillales; f__Listeriaceae; g__Listeria",
               "f0090fc348a4c45135f858d776e30dd0: \nd__Bacteria; p__Firmicutes; c__Bacilli; o__Lactobacillales; f__Lactobacillaceae; g__Lactobacillus",
               "f9eed79f2bd69b0da2385144b13658a7: \nd__Bacteria; p__Firmicutes; c__Bacilli; o__Staphylococcales; f__Staphylococcaceae",
               "5c9b89d5acf8e0020d9de19b53d21663: \nd__Bacteria; p__Firmicutes; c__Bacilli; o__Staphylococcales; f__Staphylococcaceae; g__Staphylococcus"),
      MockTaxa=c("Pseudomonas aeruginosa", "Enterococcus faecalis", "Escherichia coli", "Salmonella enterica", 
                 "Bacillus subtilis","Lactobacillus fermentum", "Listeria monocytogenes", "Lactobacillus fermentum", "Staphylococcus aureus", "Staphylococcus aureus"),
   percent=c(4,11,16,3,
             14,10,14,10,15,3)
)

reallyexpectedmock <- data_frame(
   SampleName="Expected_Mock",
   value=0,
   MockTaxa=factor(c("Pseudomonas aeruginosa", "Escherichia coli", "Salmonella enterica", "Enterococcus faecalis", 
                 "Lactobacillus fermentum","Staphylococcus aureus", "Listeria monocytogenes", "Bacillus subtilis")),
   percent=c(4.2,10.1,10.4,9.9,
            18.4,15.5,14.1,17.4))

datafullfamily %>% filter(SampleName %in% c("Mock_1","Mock_2")) %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Others", n=8)) %>% 
   mutate(TaxaOther=str_remove(TaxaOther, ".*d__Bacteria; ")) %>% 
   group_by(SampleName) %>% mutate(percent=value/sum(value)) %>% 
   ggplot(aes(x=SampleName, y=percent, fill=TaxaOther))+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=12, colour="gray20"),
         legend.position = "right", axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettess))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent family abundance",x=NULL,fill="Order")
```


```{r comtrols, fig.width=16, fig.height=20}

mockplot<- reallyexpectedmock %>% #full_join(expectedmock) %>% 
   ggplot(aes(x=SampleName, y=percent, fill=MockTaxa))+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=12, colour="gray20", margin = margin(b = 10, unit = "pt")),
         legend.position = "right",legend.direction = "vertical",
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c("#c5543a",
"#b8995a",
"#9cd250",
"#6eb386",
"#8c9fc3",
"#7e4abb",
"#c45891",
"#4b393e"))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Expected mock species abundance",x=NULL,fill="Mock species")+ 
   guides(fill = guide_legend(ncol = 2))


conreads<- datafullkingdom %>% filter(Project=="Controls") %>% 
   ggplot(aes(x=SampleName, y=value))+
   geom_col()+
   theme(legend.text = element_text(size=12, colour="gray20"),
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_y_continuous(expand=c(0,0), labels=scales::label_comma())+
   labs(y="Number of reads per sample",x=NULL,fill=NULL)

palettess2<-c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F", "#FF7F00", 
              "#CAB2D6", "#6A3D9A", "#FFFF99", "#B15928","blue1","navy",
              "maroon2","maroon4","burlywood1","burlywood4","aquamarine3","aquamarine4","grey40")

conperc<- datafullASVs %>% filter(Project=="Controls") %>% 
   left_join(read_qza(taxonomyfile)$data, by=c("ASV"="Feature.ID")) %>% 
   unite("ASVname", c("ASV", "Taxon"), sep=": \n") %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=ASVname, w=value, other_level="Others", n=20)) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>%
   group_by(SampleName) %>% mutate(percent=value/sum(value)) %>%  
   ggplot(aes(x=SampleName, y=percent, fill=str_wrap(TaxaOther, 90)))+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=12, colour="gray20", margin = margin(b = 10, unit = "pt")),
         legend.position = "right",legend.direction = "vertical",
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettess2))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent ASV abundance",x=NULL,fill="ASV")


mocklegend <- get_legend(mockplot+
   guides(fill = guide_legend(ncol = 1)))
controlpercentlegend <- get_legend(conperc+
   guides(fill = guide_legend(ncol = 1)))

plot_spacer()+conreads+mockplot+conperc+
   mocklegend + controlpercentlegend + plot_layout(guides="collect", widths=c(1,5), heights=c(2,2,5)) +
   plot_annotation(tag_levels = list(c("A","B","C","",""))) &
   theme(legend.position='none', plot.tag = element_text(face = "bold", size=20))

ggsave("LoperamideQC_mocknegativecontrols.png", width=12, height=20)

```


# Overall barplots

## Phylum level

```{r barsphylum, fig.width=16, fig.height=6}

palettep<-c("skyblue","#D95F02", "#1B9E77", "#E6AB02","#7570B3", "#66A61E","navy","maroon","gray30")


datafullphylum %>% filter(Project=="Loperamide") %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Others", n=8)) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>%
   mutate(Loperamide=recode(Loperamide, "Control water"= "Control")) %>% 
   ggplot(aes(x=SampleName, y=value, fill=TaxaOther))+
   facet_nested(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=12, colour="gray20"),
         strip.background = element_rect(color="white", fill="grey85"),
         legend.position = "bottom",axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettep))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent phyla abundance",x=NULL,fill="Phylum")+
   guides(fill = guide_legend(ncol = 3))


ggsave("LoperamideBarsPhylum.png", width=12, height=6, dpi=400)

datafullphylum  %>% 
   group_by(SampleName) %>% mutate(percent = value/sum(value)) %>% 
   group_by(name) %>% summarise(meanPhyla = mean(percent, na.rm=TRUE), sdPhyla = sd(percent, na.rm=TRUE)) %>% 
   arrange(-meanPhyla) %>% filter(meanPhyla > 0.001) %>%  gt() %>%
  fmt_percent(
    columns = c(meanPhyla, sdPhyla),
    decimals = 1
  )

```

## Order level

```{r order, fig.width=18, fig.height=6}

datafullorder %>% filter(Project=="Loperamide") %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Others", n=12)) %>% 
   mutate(TaxaOther=str_remove(TaxaOther, ".*d__Bacteria; ")) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>%
   ggplot(aes(x=SampleName, y=value, fill=TaxaOther))+
   facet_grid(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=12, colour="gray20"),
         legend.position = "bottom", axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettess))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent order abundance",x=NULL,fill="Order")
```

## Genus level

```{r genus, fig.width=16, fig.height=8}


palettess2<-c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F", "#FF7F00", 
              "#CAB2D6", "#6A3D9A", "#FFFF99", "#B15928","blue1","navy",
              "maroon2","maroon4","burlywood1","burlywood4","aquamarine3","aquamarine4","grey40")


datafullgenus %>% filter(Project=="Loperamide") %>% filter(Loperamide!="Control water") %>% 
   mutate(name=recode(name, "Unassigned; NA; NA; NA; NA; NA"="NA; NA; NA; NA; NA")) %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Others", n=21)) %>% 
   mutate(TaxaOther=str_remove(TaxaOther, ".*d__Bacteria; ")) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>%
   ggplot(aes(x=SampleName, y=value, fill=TaxaOther))+
   facet_grid(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=12, colour="gray20"),
         legend.position = "bottom", legend.direction="vertical", 
         axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettess2))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent genus abundance",x=NULL,fill="Genus")+ 
   guides(fill = guide_legend(ncol = 2))

```

```{r genuspresent, fig.width=18, fig.height=8}

palettess2<-c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F", "#FF7F00", 
              "#CAB2D6", "#6A3D9A", "#FFFF99", "#B15928","blue1","navy",
              "maroon2","maroon4","burlywood1","burlywood4","aquamarine3","aquamarine4","grey40")

datafullgenus %>% filter(Project=="Loperamide") %>% #filter(Loperamide!="Control water") %>% 
   mutate(name=recode(name, "Unassigned; NA; NA; NA; NA; NA"="NA; NA; NA; NA; NA")) %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Others", n=21)) %>% 
   mutate(TaxaOther=str_remove(TaxaOther, ".*d__Bacteria; ")) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>%
   ggplot(aes(x=SampleName, y=value, fill=TaxaOther))+
   facet_grid(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=14, colour="gray20"),
         legend.position = "bottom", legend.direction="vertical", 
         axis.text.x = element_blank(),
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettess2))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent genus abundance",x=NULL,fill="Genus")+ 
   guides(fill = guide_legend(ncol = 2))


genusbarplot <- datafullgenus %>% filter(Project=="Loperamide") %>% #filter(Loperamide!="Control water") %>% 
   mutate(name=recode(name, "Unassigned; NA; NA; NA; NA; NA"="NA; NA; NA; NA; NA")) %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Others", n=13)) %>% 
   mutate(TaxaOther=str_remove(TaxaOther, ".*d__Bacteria; ")) %>% 
   mutate(TaxaOther=recode(TaxaOther, "NA; NA; NA; NA; NA"= "Unknown bacteria")) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>%
   mutate(Loperamide=recode(Loperamide, "Control water"= "Control")) %>% 
   ggplot(aes(x=SampleName, y=value, fill=TaxaOther))+
   facet_nested(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=14, colour="gray20"),
         legend.position = "bottom", legend.direction="vertical", 
         axis.text.x = element_blank(),
         strip.background = element_rect(color="white", fill="grey85"),
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettess))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent genus abundance",x=NULL,fill="Genus")+ 
   guides(fill = guide_legend(ncol = 1))

ggsave("Loperamide16Sbars_genus.png", bg="transparent", width = 20, height = 10)



# for presentation
datafullgenus %>% filter(Project=="Loperamide") %>% filter(Loperamide!="Control water") %>% 
   mutate(name=recode(name, "Unassigned; NA; NA; NA; NA; NA"="NA; NA; NA; NA; NA")) %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Others", n=13)) %>% 
   mutate(TaxaOther=str_remove(TaxaOther, ".*d__Bacteria; ")) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>%
   ggplot(aes(x=SampleName, y=value, fill=TaxaOther))+
   facet_grid(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=13, colour="gray20"),
         legend.position = "bottom", legend.direction="vertical", strip.background = element_blank(),
         axis.text.x = element_blank(), text=element_text(family="Avenir"),
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettess))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent genus abundance",x=NULL,fill="Genus")

ggsave("Loperamide16Sbars_avenir.png", bg="transparent", width = 14, height = 9)



```

```{r genusheat, fig.width=30, fig.height=8}

datafullgenus %>% filter(Project=="Loperamide") %>% #filter(Loperamide!="Control water") %>% 
   mutate(name=recode(name, "Unassigned; NA; NA; NA; NA; NA"="NA; NA; NA; NA; NA")) %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Others", n=21)) %>% 
   mutate(TaxaOther=str_remove(TaxaOther, ".*d__Bacteria; ")) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>%
   mutate(percent=sqrt(value/sum(value))) %>% 
   ggplot(aes(x=SampleName, y=TaxaOther))+
   facet_grid(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_tile(aes(fill=percent))+
   scale_fill_gradientn(na.value = "black",labels = scales::percent,limits=c(0,0.25),
                       colors=c("white","#fecc5c","#fd8d3c","#f03b20","#bd0026","darkred")) +
   theme(legend.text = element_text(size=12, colour="gray20"),legend.key.width = unit(2,"cm"),
         legend.position = "top", axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   labs(y=NULL,x=NULL,fill="Percent genus abundance, \nsqrt transformed")


```



### Most abundant genera

```{r genusabund, fig.height=8, fig.width=4}


abundGenus <- datafullgenus %>% filter(Project=="Loperamide") %>% #filter(Loperamide!="Control water") %>% 
   mutate(name=recode(name, "Unassigned; NA; NA; NA; NA; NA"="NA; NA; NA; NA; NA")) %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Others", n=40)) %>% 
   mutate(TaxaOther=str_remove(TaxaOther, ".*d__Bacteria; ")) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>% 
   distinct(TaxaOther) %>%
   .$TaxaOther

abundGenusTopOnePercent <- datafullgenus %>% filter(Project=="Loperamide") %>% #filter(Loperamide!="Control water") %>% 
   filter(Loperamide %in% c("DMSO","Loperamide")) %>% 
   group_by(SampleName) %>% mutate(percent=(value/sum(value))) %>% 
   group_by(name) %>% mutate(meanperc=mean(percent)) %>% 
   filter(meanperc>0.01) %>% 
   mutate(GenusClean = str_replace_all(name, "; NA", "_g__"),
          GenusClean = sub(".*; ", "", GenusClean)) %>% 
   mutate(GenusClean = recode(GenusClean, "Gammaproteobacteria_g___g___g__"="Gammaproteobacteria_unknown",
                              "Unassigned_g___g___g___g___g__"="Unknown",
                              "d__Bacteria_g___g___g___g___g__"="Unknown",))

abundGenusTopOnePercent  %>% 
   ggplot(aes(x=SampleRep, y=percent, fill=GenusClean))+
   facet_grid(TimepointDay~Loperamide, scales="free", space="free")+
   geom_col(alpha=0.9)+
   theme(legend.text = element_text(size=16, colour="gray20"),
         legend.position = "bottom", legend.direction="vertical", strip.background = element_blank(),
         axis.text.x = element_blank(), panel.grid.minor.x=element_blank(),
         axis.ticks.x = element_blank())+
   scale_fill_manual(values=c(palettess, palettess))+
   scale_y_continuous(labels = scales::percent_format(), limits=c(0,1), minor_breaks = c(.25,.75), breaks=c(0,0.5,1))+
   labs(y="Percent genus abundance",x=NULL,fill="Genus  (>1% abundance)")+ 
   guides(fill = guide_legend(ncol = 1))

ggsave("LoperamideAbundGenus.png", height=8, width=4.5)
ggsave("LoperamideAbundGenus.svg", height=8, width=4.5)


```


```{r abundgenusheatmap, fig.width=10}

# use forcats lump n to get an "other genera" category
datafullgenus %>% filter(Project=="Loperamide") %>% #filter(Loperamide!="Control water") %>% 
   filter(Loperamide %in% c("DMSO","Loperamide")) %>% 
   mutate(name=recode(name, "Unassigned; NA; NA; NA; NA; NA"="Unknown")) %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=name, w=value, other_level="Sum of other genera <1%", n=12)) %>% 
   group_by(SampleName) %>% mutate(percent=(value/sum(value))) %>% 
   group_by(TaxaOther,SampleName) %>% mutate(sumpercent=sum(percent)) %>% 
   mutate(GenusClean = str_replace_all(TaxaOther, "; NA", "_g__"),
          GenusClean = sub(".*; ", "", GenusClean)) %>% 
   mutate(GenusClean = recode(GenusClean, "Gammaproteobacteria_g___g___g__"="Gammaproteobacteria_unknown",
                              "Unassigned_g___g___g___g___g__"="Unknown",
                              "d__Bacteria_g___g___g___g___g__"="Unknown")) %>%
   mutate(GenusClean=factor(GenusClean,
                            levels=c("Ensifer","Aeromonas",
                                     "Acidovorax","Comamonadaceae_g__", "Acinetobacter", "Flavobacterium",
                                     "Oxalobacteraceae_g__", "Rheinheimera",
                                     "Staphylococcus","Methylobacterium-Methylorubrum",
                                     "Gammaproteobacteria_unknown","Unknown", "Sum of other genera <1%"))) %>%
   mutate(sigGroup=case_when(GenusClean=="Ensifer" ~ "T0",
                             GenusClean=="Aeromonas" ~ "T0",
                             GenusClean=="Unknown" ~ "",
                             GenusClean=="Staphylococcus" ~ "",
                             GenusClean== "Gammaproteobacteria_unknown" ~ "",
                             GenusClean=="Methylobacterium-Methylorubrum" ~ "",
                             GenusClean=="Sum of other genera <1%" ~ "",
                             TRUE ~ "T1"),
          sigGroup=factor(sigGroup, levels=c("T0", "T1", ""))) %>%
   ggplot(aes(x=SampleName, y= fct_rev(GenusClean), fill=sqrt(sumpercent)))+
   facet_grid(sigGroup~TimepointDay+Loperamide, scales="free", space="free")+
   geom_tile()+
   theme(legend.text = element_text(size=13, colour="gray20"), strip.background = element_blank(),
         axis.text.x = element_blank(), legend.position="bottom",
         strip.text = element_text(size=13, colour="gray20"),
         axis.text.y = element_text(size=17, colour="gray20"),
         legend.title = element_text(size=15, colour="gray20"),
         panel.grid.major.x = element_blank(),panel.grid.minor.x = element_blank(),
         panel.grid.major.y = element_blank(),panel.grid.minor.y = element_blank(),
         axis.ticks.x = element_blank(), legend.key.width = unit(1,"cm"))+
 #  scale_fill_viridis_c(option="D")+
   scale_fill_gradientn(na.value = "salmon",labels = scales::percent,
                       colors=c("white","dodgerblue2","navy","black")) +
   labs(fill="Percent genus abundance\n(sqrt transformed)",x=NULL,y=NULL)


# directly from all taxa >1%
abundGenusTopOnePercent %>% 
   mutate(GenusClean=factor(GenusClean, 
                            levels=c("Ensifer","Aeromonas",
                                     "Acidovorax","Comamonadaceae_g__", "Acinetobacter", "Flavobacterium",
                                     "Oxalobacteraceae_g__", "Rheinheimera",
                                     "Gammaproteobacteria_unknown","Unknown",
                                     "Staphylococcus","Methylobacterium-Methylorubrum"))) %>% 
   mutate(sigGroup=case_when(GenusClean=="Ensifer" ~ "T0",
                             GenusClean=="Aeromonas" ~ "T0",
                             GenusClean=="Unknown" ~ "",
                             GenusClean=="Staphylococcus" ~ "",
                             GenusClean== "Gammaproteobacteria_unknown" ~ "",
                             GenusClean=="Methylobacterium-Methylorubrum" ~ "",
                             TRUE ~ "T1"),
          sigGroup=factor(sigGroup, levels=c("T0", "T1", ""))) %>% 
   ggplot(aes(x=SampleRep, y= fct_rev(GenusClean), fill=sqrt(percent)))+
   facet_grid(sigGroup~TimepointDay+Loperamide, scales="free", space="free")+
   geom_tile()+
   theme(legend.text = element_text(size=13, colour="gray20"), strip.background = element_blank(),
         axis.text.x = element_blank(), 
         axis.ticks.x = element_blank())+
   scale_fill_gradientn(na.value = "salmon",labels = scales::percent,
                       colors=c("white","darkslategrey","grey10","black")) +
   labs(fill="Percent genus\nabundance (%)",x=NULL,y=NULL)

ggsave("LoperamideAbundHeatmapGenus.png", height=5, width=10)
ggsave("LoperamideAbundHeatmapGenus.svg", height=5, width=10)

```


## ASV level

```{r asvbar, fig.width=16, fig.height=8}
datafullASVs %>% filter(Project=="Loperamide") %>% 
   left_join(read_qza(taxonomyfile)$data, by=c("ASV"="Feature.ID")) %>% 
   unite("ASVname", c("ASV", "Taxon"), sep=": ") %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=ASVname, w=value, other_level="Others", n=12)) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -value)) %>%
   ggplot(aes(x=SampleName, y=value, fill=TaxaOther))+
   facet_grid(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_col(position="fill", alpha=0.8)+
   theme(legend.text = element_text(size=12, colour="gray20"),
         legend.position = "bottom",legend.direction = "vertical", axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   scale_fill_manual(values=c(palettess))+
   scale_y_continuous(labels = scales::percent_format(), expand=c(0,0))+
   labs(y="Percent ASV abundance",x=NULL,fill="ASV")

```

And as a heatmap

```{r plotasvheat, fig.width=30, fig.height=8}
datafullASVs %>% filter(Project=="Loperamide") %>% 
   left_join(read_qza(taxonomyfile)$data, by=c("ASV"="Feature.ID")) %>% 
   unite("ASVname", c("ASV", "Taxon"), sep=": ") %>% 
   mutate(percent=sqrt(value/sum(value))) %>% 
   mutate(TaxaOther=forcats::fct_lump_n(f=ASVname, w=percent, other_level="Others", n=10)) %>% 
   mutate(TaxaOther=reorder(TaxaOther, -percent)) %>%
   ggplot(aes(x=SampleName, y=TaxaOther))+
   facet_grid(.~TimepointDay+Loperamide, scales="free", space="free")+
   geom_tile(aes(fill=percent))+
   scale_fill_gradientn(na.value = "salmon",labels = scales::percent,limits=c(0,1),
                       colors=c("white","#fecc5c","#fd8d3c","#f03b20","#bd0026","darkred")) +
   theme(legend.text = element_text(size=12, colour="gray20"),
         legend.position = "top", axis.text.x = element_blank(), 
         axis.ticks.y = element_line(inherit.blank=FALSE, color="grey30"))+
   labs(y=NULL,x=NULL,fill="Percent ASV abundance, sqrt transformed")
```

-----

# Differential Abundance

## Lefse

```{r lefse, fig.width=24, fig.height=15}


mm_lefse <- run_lefse(taxa_rank="none", norm="RLE",
    Loperamidephyseq, wilcoxon_cutoff = 0.01, group = "LoperamideTimepoint",
    kw_cutoff = 0.01, multigrp_strat = TRUE, lda_cutoff = 2)

lefseres <- marker_table(mm_lefse) %>% data.frame()

#plot_abundance(mm_lefse, group = "LoperamidePathogen")
#plot_ef_bar(mm_lefse)

lefseres %>% 
   left_join(read_qza(taxonomyfile)$data, by=c("feature"="Feature.ID")) %>% 
   mutate(Taxon=str_remove(Taxon, ".*d__Bacteria; p__")) %>% 
   unite("ASVname", c("feature", "Taxon"), sep=": ") %>% 
   filter(ef_lda > 2) %>% 
   ggplot(aes(x=ef_lda, y=reorder(ASVname,ef_lda), fill=enrich_group))+
   geom_col(position="dodge")+
   facet_grid(enrich_group~., scales="free", space="free")+
   scale_fill_manual(values=c("grey50","grey20","#a7a7ff","#0000ce", 
                              "aquamarine1", "aquamarine3","aquamarine4"))+
   theme(strip.text.y = element_blank(), legend.position="top")+
   labs(title="LefSE Results for one-against one, All p<0.01, LDA>2", 
        fill="Enriched group", x="LDA score", y=NULL)

```

## AncomBC

```{r ancom, fig.width=24, fig.height=16}

mm_ancom <- run_ancombc(Loperamidephyseq, taxa_rank="none", group = "LoperamideTimepoint", 
                        formula = "LoperamideTimepoint",norm="RLE",
                        p_adjust = "fdr", pvalue_cutoff=0.01)

ancomres <- marker_table(mm_ancom) %>% data.frame()

#plot_abundance(mm_ancom, group = "LoperamideTimepoint")
#plot_ef_bar(mm_ancom)

ancomres %>% 
   left_join(read_qza(taxonomyfile)$data, by=c("feature"="Feature.ID")) %>% 
   mutate(Taxon=str_remove(Taxon, ".*d__Bacteria; p__")) %>% 
  # unite("ASVname", c("feature", "Taxon"), sep=": ") %>% 
   filter(ef_W > 500) %>% 
   ggplot(aes(x=ef_W, y=reorder(Taxon,ef_W), fill=enrich_group))+
   geom_col(position="dodge")+
   facet_grid(enrich_group~., scales="free", space="free")+
   scale_fill_manual(values=c("black","grey80","grey50","#a7a7ff","#00006c","aquamarine3","aquamarine4"))+
   theme(strip.text.y = element_blank(), legend.position="top")+
   labs(title="ANCOM-BC results, All padj<0.01, W>100", 
        fill="Enriched group", x="Log fold change", y=NULL)

```

## Limma

### Each comparison to DMSO

```{r limma, fig.width=24, fig.height=12}


mm_limma_24h <- run_limma_voom(
    Loperamidephyseq,
    "LoperamideTimepoint",
    pvalue_cutoff = 0.01,
    taxa_rank = "none",norm="RLE",
    contrast = c("DMSO.24h", "Loperamide.24h"),
    p_adjust = "fdr"
)

mm_limma_48h <- run_limma_voom(
    Loperamidephyseq,
    "LoperamideTimepoint",
    pvalue_cutoff = 0.01,
    taxa_rank = "none",norm="RLE",
    contrast = c("DMSO.48h", "Loperamide.48h"),
    p_adjust = "fdr"
)

mm_limma_6d <- run_limma_voom(
    Loperamidephyseq,
    "LoperamideTimepoint",
    pvalue_cutoff = 0.01,
    taxa_rank = "none",norm="RLE",
    contrast = c("DMSO.6d", "Loperamide.6d"),
    p_adjust = "fdr"
)


limmares24h <- marker_table(mm_limma_24h) %>% data.frame() %>% add_column(comparison="24 h")
limmares48h <- marker_table(mm_limma_48h) %>% data.frame()%>% add_column(comparison="48 h")
limmares6d <- marker_table(mm_limma_6d) %>% data.frame() %>% add_column(comparison="6 d")

alllimmaresults <- full_join(limmares24h, limmares48h)  %>% full_join(limmares6d)

alllimmaresults %>% 
   left_join(read_qza(taxonomyfile)$data, by=c("feature"="Feature.ID")) %>% 
   mutate(Taxon=str_remove(Taxon, ".*d__Bacteria; p__")) %>% 
   unite("ASVname", c("feature", "Taxon"), sep=": ") %>% 
   ggplot(aes(x=ef_logFC, y=reorder(ASVname,ef_logFC), fill=enrich_group))+
   geom_col(position="dodge")+
   facet_grid(comparison~., scales="free", space="free")+
   scale_fill_manual(values=c("#0000ce", "#00006c","aquamarine1", "aquamarine3","aquamarine4"))+
   theme(strip.text.y = element_blank(), legend.position="top")+
   labs(title="Limma Results, each comparison separately. All padj<0.01", 
        fill="Enriched group", x="Log fold change", y=NULL)
   

alllimmaresults %>% 
   left_join(read_qza(taxonomyfile)$data, by=c("feature"="Feature.ID")) %>% 
   mutate(Taxon=str_remove(Taxon, ".*d__Bacteria; p__")) %>% 
   unite("ASVname", c("feature", "Taxon"), sep=": ") %>% 
   filter(abs(ef_logFC) > 5) %>% 
   ggplot(aes(x=ef_logFC, y=reorder(ASVname,ef_logFC), fill=enrich_group))+
   geom_col(position="dodge")+
   facet_grid(comparison~., scales="free", space="free")+
   scale_fill_manual(values=c("#0000ce", "#00006c", "aquamarine3","aquamarine4"))+
   theme(legend.position="top")+
   labs(title="Limma Results at ASV level, each comparison separately. All padj<0.01 and fold change >|5|", 
        fill="Enriched group", x="Log fold change", y=NULL)

```

### All groups together

```{r limmaall, fig.width=26, fig.height=7}

mm_limma_all <- run_limma_voom(
    Loperamidephyseq,norm="RLE",
    "LoperamideTimepoint",
    pvalue_cutoff = 0.01,
    taxa_rank = "none",
    p_adjust = "fdr"
)

#plot_ef_bar(mm_limma)

limmares_all <- marker_table(mm_limma_all) %>% data.frame() %>% add_column(comparison="All")


limmares_all %>% 
   left_join(read_qza(taxonomyfile)$data, by=c("feature"="Feature.ID")) %>% 
   mutate(Taxon=str_remove(Taxon, ".*d__Bacteria; p__")) %>% 
   unite("ASVname", c("feature", "Taxon"), sep=": ") %>% 
   filter(ef_F_statistic > 300) %>% 
   ggplot(aes(x=ef_F_statistic, y=reorder(ASVname,ef_F_statistic),
              fill=enrich_group))+
   geom_col(position="dodge")+
   facet_grid(comparison~., scales="free", space="free")+
   scale_fill_manual(values=c("#a7a7ff","#0000ce", "aquamarine1", "aquamarine3","aquamarine4"))+
   theme(legend.position="top")+
   labs(title="Limma Results, all comparisons. All padj<0.01 and ef_F_statistic > 1000", 
        fill="Enriched group", x="Log fold change", y=NULL)

```

### At genus level

```{r genuslimma, fig.width=10, fig.height=10}

mm_limma_24h <- run_limma_voom(
    Loperamidephyseq, norm="RLE",
    "LoperamideTimepoint",
    pvalue_cutoff = 0.05,
    taxa_rank = "Genus",
    contrast = c("DMSO.24h", "Loperamide.24h"),
    p_adjust = "fdr"
)

mm_limma_48h <- run_limma_voom(
    Loperamidephyseq,  norm="RLE",
    "LoperamideTimepoint",
    pvalue_cutoff = 0.05,
    taxa_rank = "Genus",
    contrast = c("DMSO.48h", "Loperamide.48h"),
    p_adjust = "fdr"
)

mm_limma_6d <- run_limma_voom(
    Loperamidephyseq,  norm="RLE",
    "LoperamideTimepoint",
    pvalue_cutoff = 0.05,
    taxa_rank = "Genus",
    contrast = c("DMSO.6d", "Loperamide.6d"),
    p_adjust = "fdr"
)


limmares24h <- marker_table(mm_limma_24h) %>% data.frame() %>% add_column(comparison="T0")
limmares48h <- marker_table(mm_limma_48h) %>% data.frame()%>% add_column(comparison="T1")
limmares6d <- marker_table(mm_limma_6d) %>% data.frame() %>% add_column(comparison="T5")

alllimmaresults <- full_join(limmares24h, limmares48h)  %>% full_join(limmares6d)


alllimmaresults %>% 
   filter(abs(ef_logFC) > 2) %>% 
   ggplot(aes(x=ef_logFC, y=reorder(feature,ef_logFC), fill=enrich_group))+
   geom_col(position="dodge")+
   facet_grid(comparison~., scales="free", space="free")+
   scale_fill_manual(values=c('#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759'))+
   theme(legend.text = element_text(size=14, colour="gray20"),
         panel.grid.major.y = element_blank(),
         legend.position = "top", legend.direction="vertical", strip.background = element_blank(),
         text=element_text(family="Avenir"))+
   labs(fill="Enriched group", x="Log fold change", y=NULL)+ 
   guides(fill = guide_legend(ncol = 2))

ggsave("LoperamideLimmaGenus_avenir.png", bg="transparent", width = 7, height = 10)

```



#### Limma plot with abundant highlighed

```{r limmagenusabund, fig.height=9, fig.width=8}

limmagenus<- alllimmaresults %>% 
   mutate(enrich_group = factor(enrich_group, 
                          levels = c("DMSO.24h", "DMSO.48h","DMSO.6d",
                                     "Loperamide.24h","Loperamide.48h","Loperamide.6d"),
                          labels = c("DMSO.T0", "DMSO.T1","DMSO.T5",
                                     "Loperamide.T0","Loperamide.T1","Loperamide.T5"))) %>% 
   mutate(feature = recode(feature, "Actinobacteriota_c___o___f___g__" = "Actinobacteriota")) %>% 
   mutate(highlighttaxa = case_when(feature %in% abundGenusTopOnePercent$GenusClean ~ ">1% abundance",
                                    TRUE ~ "low abundance")) %>% 
   mutate(featurebold = case_when(feature %in% abundGenusTopOnePercent$GenusClean ~ 
                                     paste0("<span style='color:black'>**",feature,"**</span>"),
                                    TRUE ~ feature)) %>% 
   filter(abs(ef_logFC) > 2) %>% 
   ggplot(aes(x=ef_logFC, y=reorder(featurebold,ef_logFC), fill=enrich_group))+
   geom_col(position="dodge", aes(color=highlighttaxa), lwd=1.5, width=0.8)+
   facet_grid(comparison~., scales="free", space="free")+
   scale_fill_manual(values=c('#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759'))+
   scale_color_manual(values=c("black", "transparent"), guide = "none") +
   theme(legend.position="bottom",legend.direction = "vertical",
         axis.text.y=element_markdown(color="grey40"),
         panel.grid.major.x = element_line(color="gray70"),panel.grid.major.y = element_line(size=0.2))+
   labs(#title="Limma Results at genus level, each comparison separately. 
               #All padj<0.05 and fold change >|2|", 
        fill="Enriched \ngroup", x="Log fold change", y=NULL)+ 
   guides(fill = guide_legend(ncol = 2,title.position = "left", hjust = 1))
limmagenus
ggsave("LoperamideLimmaGenus.png", bg="transparent", width = 7, height = 10)
```


#### Limma plot only abundant taxa

```{r limmagenusabundenrich, fig.width=5 }

abundGenusC <- abundGenusTopOnePercent %>% 
   distinct(name) %>%
   .$name %>% str_replace_all(., "; NA", "_g__") %>% sub(".*; ", "", .)

alllimmaresultssigabund<- alllimmaresults %>% 
   mutate(enrich_group = factor(enrich_group, 
                          levels = c("DMSO.24h", "DMSO.48h","DMSO.6d",
                                     "Loperamide.24h","Loperamide.48h","Loperamide.6d"),
                          labels = c("DMSO.T0", "DMSO.T1","DMSO.T5",
                                     "Loperamide.T0","Loperamide.T1","Loperamide.T5"))) %>% 
   mutate(feature = recode(feature, "Actinobacteriota_c___o___f___g__" = "Actinobacteriota")) %>% 
   filter(feature %in% abundGenusC) 

alllimmaresultssigabund %>% 
   ggplot(aes(x=ef_logFC, y=reorder(feature,-ef_logFC), fill=enrich_group))+
   geom_col(position="dodge")+
   facet_grid(comparison~., scales="free", space="free")+
   scale_fill_manual(values=c('#1c5580','#1c5580', '#0fc08e','#0fc08e'))+
   theme(legend.position="none",
         panel.grid.major.x = element_line(color="gray70", size=0.5),
         panel.grid.major.y = element_blank())+
   labs(#title="Limma Results at genus level, each comparison separately. 
               #All padj<0.05 and fold change >|2|", 
        fill="Enriched \ngroup", x="Log fold change", y=NULL)+ 
   guides(fill = guide_legend(ncol = 2,title.position = "left", hjust = 1))


ggsave("LoperamideLimmaGenusAbund.svg", bg="transparent", width = 4, height = 2.5)




```


-----

# Venn Diagrams

## Genus level


```{r venngenus, fig.width=10, fig.height=4}

library(UpSetR)
library(gplots)
library(ComplexHeatmap)

DMSO24hsetG <- datafullgenus %>% filter(LoperamideTimepoint=="DMSO.T0") %>% 
   filter(value >5) %>% distinct(name) %>% .$name
DMSO48hsetG <- datafullgenus %>% filter(LoperamideTimepoint=="DMSO.T1") %>% 
   filter(value >5) %>% distinct(name) %>% .$name
DMSO6dsetG <- datafullgenus %>% filter(LoperamideTimepoint=="DMSO.T5") %>% 
   filter(value >5) %>% distinct(name) %>% .$name
Loperamide24hsetG <- datafullgenus %>% filter(LoperamideTimepoint=="Loperamide.T0") %>% 
   filter(value >5) %>% distinct(name) %>% .$name
Loperamide48hsetG <- datafullgenus %>% filter(LoperamideTimepoint=="Loperamide.T1") %>% 
   filter(value >5) %>% distinct(name) %>% .$name
Loperamide6dsetG <- datafullgenus %>% filter(LoperamideTimepoint=="Loperamide.T5") %>% 
   filter(value >5) %>% distinct(name) %>% .$name

read_setsG = list("DMSO T0" = DMSO24hsetG, "DMSO T1" = DMSO48hsetG, "DMSO T5" = DMSO6dsetG, 
                 "Loperamide T0" = Loperamide24hsetG, "Loperamide T1" = Loperamide48hsetG,
                 "Loperamide T5" = Loperamide6dsetG)
mG = make_comb_mat(read_setsG)

(upsetGplot <- UpSet(mG,      
      set_order = order(c("DMSO.T0","DMSO.T1","DMSO.T5",
                          "Loperamide.T0","Loperamide.T1","Loperamide.T5")),
      pt_size = unit(.5, "cm"),lwd=2.5,
      comb_col = c("black","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50",
                   "grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50",
                   "grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","maroon",
                   "grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50",
                   "maroon","grey50","grey50",
                   '#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759'),
      right_annotation = rowAnnotation(" " = anno_barplot(set_size(mG), bar_width=0.7,
                                                         axis_param = list(side = "top",labels_rot = 0),
                                                         border = FALSE, annotation_name_side = "top", 
                                                         gp = gpar(fill = c('#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759')),
                                                         width = unit(4, "cm"))),
      row_names_side = "left",
      top_annotation = upset_top_annotation(mG,bar_width = 0.9, height = unit(6, "cm"))))

```

### Figure with barplots


```{r genussummary, fig.width=15, fig.height=20}

genusbarplot / ggplotify::as.grob(upsetGplot)+plot_annotation(tag_levels = "A")+plot_layout(heights=c(3,2.5)) &
   theme(plot.tag = element_text(face = "bold", size=20))

ggsave("LoperamideBarsGenusUpset.png", width=12.5, height=14, dpi=400)

```




## ASV level

```{r vennasv, fig.width=12, fig.height=4}

library(UpSetR)
library(gplots)
library(ComplexHeatmap)

DMSO24hset <- datafullASVs %>% filter(LoperamideTimepoint=="DMSO.T0") %>% 
   filter(value >5) %>% distinct(ASV) %>% .$ASV
DMSO48hset <- datafullASVs %>% filter(LoperamideTimepoint=="DMSO.T1") %>% 
   filter(value >5) %>% distinct(ASV) %>% .$ASV
DMSO6dset <- datafullASVs %>% filter(LoperamideTimepoint=="DMSO.T5") %>% 
   filter(value >5) %>% distinct(ASV) %>% .$ASV
Loperamide24hset <- datafullASVs %>% filter(LoperamideTimepoint=="Loperamide.T0") %>% 
   filter(value >5) %>% distinct(ASV) %>% .$ASV
Loperamide48hset <- datafullASVs %>% filter(LoperamideTimepoint=="Loperamide.T1") %>% 
   filter(value >5) %>% distinct(ASV) %>% .$ASV
Loperamide6dset <- datafullASVs %>% filter(LoperamideTimepoint=="Loperamide.T5") %>% 
   filter(value >5) %>% distinct(ASV) %>% .$ASV

read_sets = list("DMSO 24h" = DMSO24hset, "DMSO 48h" = DMSO48hset, "DMSO 6d" = DMSO6dset, 
                 "Loperamide 24h" = Loperamide24hset, "Loperamide 48h" = Loperamide48hset,
                 "Loperamide 6d" = Loperamide6dset)
m = make_comb_mat(read_sets)

UpSet(m,      
      set_order = order(c("DMSO.24h","DMSO.48h","DMSO.6d", "Loperamide.24h","Loperamide.48h","Loperamide.6d")),
      pt_size = unit(.5, "cm"),lwd=2.5,
      comb_col = c("black","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50",
                   "grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50",
                   "grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50","grey50",
                   "grey50","grey50","grey50","grey50","maroon","grey50","grey50","grey50","grey50","grey50","grey50","grey50",
                   "grey50","grey50","grey50","maroon","maroon", "grey50","grey50","grey50", 
                   "#a7a7ff","#0000ce", "#00006c","aquamarine1", "aquamarine3","aquamarine4"),
      right_annotation = rowAnnotation(" " = anno_barplot(set_size(m), bar_width=0.7,
                                                         axis_param = list(side = "top",labels_rot = 0),
                                                         border = FALSE, annotation_name_side = "top",
                                                         gp = gpar(fill = c("#a7a7ff","#0000ce", "#00006c",
                                                                            "aquamarine1", "aquamarine3","aquamarine4")),
                                                         width = unit(4, "cm"))),
      row_names_side = "left",
      top_annotation = upset_top_annotation(m,bar_width = 0.9, height = unit(6, "cm"))) 

```


```{r venn2, fig.width=10}

library(UpSetR)
library(gplots)
library(ComplexHeatmap)

Con24hset <- datafullASVs %>% filter(LoperamideTimepoint=="ControlWater.T0") %>% 
   filter(value >5) %>% distinct(ASV) %>% .$ASV
Con48hset <- datafullASVs %>% filter(LoperamideTimepoint=="ControlWater.T1") %>% 
   filter(value >5) %>% distinct(ASV) %>% .$ASV
Con6dset <- datafullASVs %>% filter(LoperamideTimepoint=="ControlWater.T5") %>% 
   filter(value >5) %>% distinct(ASV) %>% .$ASV


read_sets = list("Water 24h" = Con24hset, "Water 48h" = Con48hset,
                 "Water 6d" = Con6dset,
                 "DMSO 24h" = DMSO24hset, "DMSO 48h" = DMSO48hset, "DMSO 6d" = DMSO6dset, 
                 "Loperamide 24h" = Loperamide24hset, "Loperamide 48h" = Loperamide48hset,
                 "Loperamide 6d" = Loperamide6dset)
m2 = make_comb_mat(read_sets)
m2sub = m2[comb_size(m2) >= 10]

UpSet(m2sub,
      set_order = order(c("ControlWater.24h","ControlWater.48h","ControlWater.6d",
                          "DMSO.24h","DMSO.48h","DMSO.6d", 
                          "Loperamide.24h","Loperamide.48h","Loperamide.6d")),
      pt_size = unit(.5, "cm"),lwd=3,
      comb_col = c("black","grey","grey","grey","grey","grey","grey","grey","maroon","grey","grey","grey",
                   "grey","grey","grey","grey","grey","grey","grey","grey","grey","grey","grey","grey",
                   "grey","grey","grey","grey","grey","grey","grey","grey","grey","grey","grey","maroon",
                   "grey","grey","grey","grey","grey","grey","maroon","maroon","grey",
                   "grey80","grey50","grey20","#a7a7ff","#0000ce",
                   "#00006c","aquamarine1", "aquamarine3","aquamarine4"),
      right_annotation = rowAnnotation(" " = anno_barplot(set_size(m2sub), bar_width=0.7,
                                                         axis_param = list(side = "top",labels_rot = 0),
                                                         border = FALSE, annotation_name_side = "top",
                                                         gp = gpar(fill = c("grey80","grey50","grey20","#a7a7ff","#0000ce",
                                                                            "#00006c","aquamarine1", "aquamarine3","aquamarine4")),
                                                         width = unit(4, "cm"))),
      row_names_side = "left",
      top_annotation = upset_top_annotation(m2sub,bar_width = 0.9, height = unit(6, "cm"))) 

```





-----

# Alpha Diversity

## Richness

```{r alpharich, fig.width=10, fig.height=6}

# calculate diversity
datafullASVs %>% 
   group_by(SampleID, LoperamideTimepoint, TimepointDay, Loperamide) %>% 
   filter(value>0) %>% 
   filter(Project=="Loperamide") %>% 
   count() %>% 
   ggplot(aes(x=Loperamide,y=n, fill=Loperamide))+
   geom_jitter(width=0.15, size=3, shape=23, alpha=0.8)+
   geom_boxplot(alpha=0.8)+
   facet_grid(.~TimepointDay, scales="free", space="free")+
   stat_compare_means(ref.group="DMSO", label = "p.format", vjust=-0.5)+
   labs(title="Richness (number of ASVs per sample)", caption="(n=5). Wilcoxon Test, relative to 24 h DMSO condition",
        x=NULL,y="Number of ASVs",fill="LoperamideTimepoint")+
   scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e')) +
   scale_fill_manual(values=c('#000000', '#1c5580', '#0fc08e')) +
   theme_bw()+
   theme(legend.position = "none", strip.background = element_blank(), strip.text=element_text(size=16))
 

```

## Richness chao

```{r alphachao, fig.width=10, fig.height=6}

chao <- estimateR(matrixLoperamide) %>% t() %>% as.data.frame()
metadataLoperamide$chao <- chao$S.chao1

# calculate diversity
metadataLoperamide %>% filter(Project=="Loperamide") %>% 
   ggplot(aes(x=Loperamide,y=chao, fill=Loperamide))+
   geom_jitter(width=0.15, size=3, shape=23, alpha=0.8)+
   geom_boxplot(alpha=0.8)+
      facet_grid(.~TimepointDay, scales="free", space="free")+
   stat_compare_means(ref.group="DMSO", label = "p.format", vjust=-0.5)+
   labs(title="Chao Index", caption="(n=5). Wilcoxon Test, relative to 24 h DMSO condition",
        x=NULL,y="Chao Index",fill="Loperamide Timepoint")+
   scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e')) +
   scale_fill_manual(values=c('#000000', '#1c5580', '#0fc08e')) +
   theme_bw()+
   theme(legend.position = "none", strip.background = element_blank(), strip.text=element_text(size=16))
 

```

## Simpsons and Shannon

```{r alphasimpshan, fig.width=14, fig.height=5}

#SVsLoperamide <- select(as.data.frame(SVs), metadataLoperamide$SampleID) %>% rownames_to_column("ASV") %>% 
#      mutate_if(is.numeric, scale2) %>% 
#   filter(!ASV %in% eukASVs) %>% column_to_rownames(var="ASV") %>% t()

# calculate diversity
diversityshannon<-diversity(matrixLoperamide, index="shannon")
diversitysimpsons<-diversity(matrixLoperamide, index="simpson")
# add into metadata variable
metadataLoperamide$Simpsons<-diversitysimpsons
metadataLoperamide$Shannon<-diversityshannon

# diversity plots
metadataLoperamide %>% filter(Project=="Loperamide") %>% 
   pivot_longer(Simpsons:Shannon, names_to="DiversityIndex", values_to="Value") %>% 
   ggplot(aes(x=LoperamideTimepoint,y=Value, fill=LoperamideTimepoint))+
   geom_jitter(width=0.15, size=3, shape=23, alpha=0.8)+
   geom_boxplot(alpha=0.8)+
   facet_wrap(~DiversityIndex, scales="free")+
   stat_compare_means(ref.group="DMSO.24h", label = "p.format", vjust=-0.5)+
   labs(caption="(n=5). Wilcoxon Test, relative to 24 h DMSO condition", x=NULL,
        y="Shannon or Simpson's Index of Diversity",fill="LoperamideTimepoint")+
   scale_color_manual(values=c("grey80","grey50","grey20","#a7a7ff","#0000ce", "#00006c","aquamarine1", "aquamarine3","aquamarine4")) +
   scale_fill_manual(values=c("grey80","grey50","grey20","#a7a7ff","#0000ce", "#00006c","aquamarine1", "aquamarine3","aquamarine4")) +
   theme_bw()+
   theme(legend.position = "none", strip.background = element_blank(), strip.text=element_text(size=16))
```
 
```{r alphasimp, fig.width=7, fig.height=4}

# just simpsons
metadataLoperamide %>% filter(Project=="Loperamide") %>% 
   ggplot(aes(x=Loperamide,y=Simpsons, fill=Loperamide))+
   geom_jitter(width=0.15, size=3, shape=23, alpha=0.8)+
   geom_boxplot(alpha=0.8)+
   facet_wrap(~TimepointDay, scales="free_x")+
   stat_compare_means(ref.group="DMSO", label = "p.format", vjust=-0.5, size=2.5)+
   labs(caption="(n=5). Wilcoxon Test, relative to DMSO condition", x=NULL,
        y="Simpson's Index of Diversity",fill="LoperamideTimepoint")+
   scale_fill_manual(values=c('#000000', '#1c5580', '#0fc08e')) +
   scale_y_continuous(limits=c(NA,1))+
   theme_bw()+
   theme(legend.position = "none", strip.background = element_blank(), strip.text=element_text(size=16))

```


```{r alphasimppresent, fig.width=10, fig.height=6}


statsSimpsons <- compare_means(data=metadataLoperamide, 
                            Simpsons~Loperamide, 
                            group.by = c("TimepointDay")) %>% 
   filter(p.format<0.05 & group1=="DMSO")



convalpha<- metadataLoperamide %>% filter(Project=="Loperamide") %>% 
   mutate(Loperamide=factor(Loperamide, levels=c("Control water", "DMSO", "Loperamide"),
                            labels=c("Control", "DMSO", "Loperamide"))) %>% 
   ggplot(aes(x=Loperamide,y=Simpsons,shape=Loperamide, fill=Loperamide)) +
   geom_boxplot(aes(), alpha=0.8)+
   geom_jitter(aes(), width=0.1, size=3, alpha=0.8)+
   geom_hline(yintercept=0)+
   geom_text(data=statsSimpsons, aes(label="*", x=group2, y=0.9, fill=NA, shape=NA), size=9, color="#0fc08e",
            show.legend=FALSE)+
   facet_grid(~TimepointDay, scales="free_x")+
   labs( x=NULL, y="Simpson's Index of Diversity\nConventional fish",fill=NULL, shape=NULL)+
   scale_fill_manual(values=c('#000000', '#1c5580', '#0fc08e')) +
   scale_shape_manual(values=c(21,24,22))+
   scale_y_continuous(limits=c(0,1.0), expand=c(0,0))+
   theme(text=element_text(size=16), 
       #  axis.ticks.x = element_blank(),axis.text.x = element_blank(),
         strip.text=element_text(size=20),
         legend.position = "none", strip.background = element_blank(),
         panel.background = element_rect(color="grey80", size=0.8),
         legend.text = element_text(size=16))
convalpha   

ggsave("LoperamideAlphaDiv16S.png", width=12, height=4, dpi=400)

convalpha2 <- metadataLoperamide %>% filter(Project=="Loperamide") %>% 
   mutate(Loperamide=factor(Loperamide, levels=c("Control water", "DMSO", "Loperamide"),
                            labels=c("Control", "DMSO", "Loperamide"))) %>% 
   ggplot(aes(x=TimepointDay, y=Simpsons, 
             fill=Loperamide, color=Loperamide, shape=Loperamide))+
   geom_boxplot(color="black", alpha=0.8, show.legend = FALSE)+
   geom_point(size=2, position=position_jitterdodge(jitter.width=0.25)) +
   geom_text(data=statsSimpsons, aes(label="*", x=TimepointDay, 
                                        y=0.9, fill=NA, shape=NA), 
             size=9, color="#0fc08e", show.legend=FALSE, nudge_x=.3)+
   scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
   scale_fill_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
   scale_y_continuous(limits=c(0,1.0), expand=c(0,0))+
   theme(legend.position = "none")+
   labs(x=NULL, y="Simpson's Index of Diversity",fill=NULL, shape=NULL, color=NULL,
        title="Conventional fish")+
   theme(legend.position = "bottom", plot.title = element_text(size=20),
         panel.background = element_rect(color="grey80", size=0.8))+
   guides(color=guide_legend(override.aes = list(size=4)))




```



-----



<!-- ## All conditions -->

<!-- ```{r loperamidebetaalltogether, fig.show='hide', results=FALSE, echo=FALSE} -->

<!-- library(phyloseq) -->

<!-- Loperamidephyseq <- phyloseq::subset_samples(physeq, Project=="Loperamide") %>%  -->
<!--    phyloseq::subset_taxa((Kingdom!="d__Eukaryota") | is.na(Kingdom)) %>%  -->
<!--    phyloseq::subset_taxa((Kingdom!="d__Archaea") | is.na(Kingdom)) %>%  -->
<!--    phyloseq::subset_taxa((Phylum!="Cyanobacteria") | is.na(Phylum)) -->

<!-- raremax <- min(sample_sums(Loperamidephyseq)) -->
<!-- ps.rarefied <- rarefy_even_depth(Loperamidephyseq, rngseed=1, sample.size=raremax, replace=F) -->
<!-- ps.rarefied -->

<!-- ordination_NMDS_bray <- ordinate(ps.rarefied, method="NMDS", distance="bray", scale=TRUE) -->

<!-- plot_ordination(ps.rarefied, ordination_NMDS_bray, color="LoperamideTimepoint") +  -->
<!--   geom_point(size=4, alpha=0.8, aes(shape=TimepointDay)) +  -->
<!--    ggtitle("NMDS Bray-Curtis, all Samples") +  -->
<!--   stat_ellipse(type = "norm", level=0.95, aes(linetype = LoperamideTimepoint)) + -->
<!--   #annotate(geom="text", x=0.9, y=-2, label=paste("Stress = ",round(ordination_NMDS_bray$stress,4)), color="black") + -->
<!--   scale_color_manual(values=c("grey80","grey50","grey20","#a7a7ff","#0000ce", "#00006c","aquamarine1", "aquamarine3","aquamarine4")) + -->
<!--   scale_fill_manual(values=c("grey80","grey50","grey20","#a7a7ff","#0000ce", "#00006c","aquamarine1", "aquamarine3","aquamarine4")) + -->
<!--   scale_linetype_manual(values=c("solid","solid","solid","dotted","dotted","dotted","dashed","dashed","dashed"))+ -->
<!--   theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right", -->
<!--         legend.title = element_blank(),legend.box="horizontal") -->

<!-- ``` -->


<!-- ## Just DMSO and loperamide -->


<!-- ```{r loperamidedmsobeta} -->

<!-- ps.rarefied.nowat <- ps.rarefied %>%  -->
<!--    phyloseq::subset_samples(LoperamidePathogen=="DMSO 1:100" |  -->
<!--                                LoperamidePathogen=="Loperamide 10mg/L")  -->

<!-- ordination_NMDS_bray_nowat = ordinate(ps.rarefied.nowat, method="NMDS", distance="bray", scale=TRUE) -->

<!-- plot_ordination(ps.rarefied.nowat, ordination_NMDS_bray_nowat, color="LoperamideTimepoint") +  -->
<!--   geom_point(size=4, alpha=0.8, aes(shape=TimepointDay, fill=LoperamideTimepoint)) +  -->
<!--   stat_ellipse(type = "norm", level=0.95, aes(linetype = LoperamideTimepoint)) + -->
<!--   annotate(geom="text", x=0.9, y=-2, label=paste("Stress = ",round(ordination_NMDS_bray_nowat$stress,4)), color="black") + -->
<!--    scale_color_manual(values=c("#a7a7ff","#0000ce", "#00006c","aquamarine1", "aquamarine3","aquamarine4")) + -->
<!--    scale_fill_manual(values=c("#a7a7ff","#0000ce", "#00006c","aquamarine1", "aquamarine3","aquamarine4")) + -->
<!--    scale_shape_manual(values = c(21,22,23,21,22,23))+ -->
<!--    scale_linetype_manual(values=c("solid","dotted","dashed","solid","dotted","dashed"))+ -->
<!--    theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "bottom", -->
<!--         legend.title = element_blank(),legend.box="verical",legend.key.width = unit(2,"cm"))  -->
<!-- #   guides(shape = guide_legend(ncol = 2)) -->


<!-- ``` -->


<!-- ## Each timepoint separately -->


<!-- ```{r loperamidebeta, fig.show='hide', results=FALSE, echo=FALSE} -->

<!-- ps.rarefied.1 <- ps.rarefied %>% phyloseq::subset_samples(Timepoint==1)  -->
<!-- ordination_NMDS_bray_1 = ordinate(ps.rarefied.1, method="NMDS", distance="bray", scale=TRUE) -->

<!-- loperamide1 <- plot_ordination(ps.rarefied.1, ordination_NMDS_bray_1, color="LoperamidePathogen") +  -->
<!--   geom_point(size=4, alpha=0.8, aes(shape=LoperamidePathogen, fill=LoperamidePathogen)) +  -->
<!--   stat_ellipse(type = "norm", level=0.95, aes(linetype = LoperamidePathogen)) + -->
<!--   annotate(geom="text", x=0.9, y=-1, label=paste("Stress = ",round(ordination_NMDS_bray$stress,4)), color="black") + -->
<!--   scale_fill_manual(values=c("grey20","navy","aquamarine4"))+ -->
<!--    scale_colour_manual(values=c("grey20","navy","aquamarine4"))+ -->
<!--    scale_shape_manual(values = c(21,22,23))+ -->
<!--    scale_linetype_manual(values=c("solid","dotted","dashed","solid","dotted","dashed"))+ -->
<!--   theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right", -->
<!--         legend.title = element_blank(),legend.box="horizontal")+ggtitle("Timepoint 1. 24h") -->

<!-- ### -->

<!-- ps.rarefied.2 <- ps.rarefied %>% phyloseq::subset_samples(Timepoint==2)  -->
<!-- ordination_NMDS_bray_2 = ordinate(ps.rarefied.2, method="NMDS", distance="bray", scale=TRUE) -->

<!-- loperamide2 <- plot_ordination(ps.rarefied.2, ordination_NMDS_bray_2, color="LoperamidePathogen") +  -->
<!--   geom_point(size=4, alpha=0.8, aes(shape=LoperamidePathogen, fill=LoperamidePathogen)) +  -->
<!--   stat_ellipse(type = "norm", level=0.95, aes(linetype = LoperamidePathogen)) + -->
<!--   annotate(geom="text", x=0.9, y=-1, label=paste("Stress = ",round(ordination_NMDS_bray$stress,4)), color="black") + -->
<!--   scale_fill_manual(values=c("grey20","navy","aquamarine4"))+ -->
<!--    scale_colour_manual(values=c("grey20","navy","aquamarine4"))+ -->
<!--    scale_shape_manual(values = c(21,22,23))+ -->
<!--    scale_linetype_manual(values=c("solid","dotted","dashed","solid","dotted","dashed"))+ -->
<!--   theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right", -->
<!--         legend.title = element_blank(),legend.box="horizontal")+ggtitle("Timepoint 2. 48h") -->

<!-- ### -->

<!-- ps.rarefied.3 <- ps.rarefied %>% phyloseq::subset_samples(Timepoint==3)  -->
<!-- ordination_NMDS_bray_3 = ordinate(ps.rarefied.3, method="NMDS", distance="bray", scale=TRUE) -->

<!-- loperamide3 <- plot_ordination(ps.rarefied.3, ordination_NMDS_bray_3, color="LoperamidePathogen") +  -->
<!--   geom_point(size=4, alpha=0.8, aes(shape=LoperamidePathogen, fill=LoperamidePathogen)) +  -->
<!--   stat_ellipse(type = "norm", level=0.95, aes(linetype = LoperamidePathogen)) + -->
<!--   annotate(geom="text", x=0.9, y=-1, label=paste("Stress = ",round(ordination_NMDS_bray$stress,4)), color="black") + -->
<!--   scale_fill_manual(values=c("grey20","navy","aquamarine4"))+ -->
<!--    scale_colour_manual(values=c("grey20","navy","aquamarine4"))+ -->
<!--    scale_shape_manual(values = c(21,22,23))+ -->
<!--    scale_linetype_manual(values=c("solid","dotted","dashed","solid","dotted","dashed"))+ -->
<!--   theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right", -->
<!--         legend.title = element_blank(),legend.box="horizontal")+ggtitle("Timepoint 3. 6d") -->

<!-- ``` -->

<!-- ```{r betaallplot1,  fig.width=12, fig.height=6} -->

<!-- library(patchwork) -->
<!-- loperamide1+loperamide2+loperamide3 + plot_layout(guides = "collect") &  -->
<!--    theme(legend.position = "bottom", text=element_text(family="Avenir"), -->
<!--          plot.background = element_blank(), legend.background = element_blank()) & coord_fixed() -->
<!-- #ggsave("LoperamideBetaAllSeparate_avenir.png", bg="transparent", width = 12, height = 6) -->

<!-- loperamide1 +loperamide2 + loperamide3 + guide_area() + plot_layout(guides = "collect") &  -->
<!--    theme(legend.position = "right", text=element_text(family="Avenir"), -->
<!--          plot.background = element_blank(), legend.background = element_blank()) -->
<!-- #ggsave("LoperamideBetaAllSeparate3_avenir.png", bg="transparent", width = 7, height = 6) -->

<!-- ``` -->

<!-- ## Each timepoint separately, just DMSO and loperamide -->


<!-- ```{r loperamidebetadmsolop,fig.show='hide', results=FALSE, echo=FALSE} -->

<!-- metadataLoperamide1sub <- metadataLoperamide1  %>% filter(Loperamide!="Control water") -->
<!-- SVsLoperamidesub1<- matrixLoperamidePercent[c(6:15),] -->
<!-- adonissub1<-adonis2(SVsLoperamidesub1~Loperamide, data=metadataLoperamide1sub, by=NULL, method="bray", k=2) -->

<!-- sol<-metaMDS(SVsLoperamidesub1,distance = "bray", k = 2, trymax = 200) -->
<!-- sol$stress; stressplot(sol) -->

<!-- NMDS=data.frame(x=sol$point[,1],y=sol$point[,2], -->
<!--                 LoperamideTreatment=as.factor(metadataLoperamide1sub$Loperamide), -->
<!--                 TimepointDay=as.factor(metadataLoperamide1sub$TimepointDay), -->
<!--                 Rep=as.factor(metadataLoperamide1sub$SampleRep)) -->
<!-- plot.new() -->
<!-- ord<-ordiellipse(sol, NMDS$LoperamideTreatment, display = "sites", kind ="sd", conf = 0.95, label = T) -->
<!-- dev.off() -->
<!-- df_ell <- data.frame() -->
<!-- for(g in levels(NMDS$LoperamideTreatment)){ -->
<!--   if(g!="" && (g %in% names(ord))){ -->
<!--     df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTreatment==g,], -->
<!--                                                      veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTreatment=g))}} -->
<!-- head(df_ell) -->
<!-- NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTreatment),mean) -->
<!-- loperamide1sub <- ggplot(data=NMDS,aes(x,y))+ -->
<!--   geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTreatment,colour=LoperamideTreatment), size=1) + -->
<!--   geom_point(size=4, alpha=0.9,aes(shape=LoperamideTreatment,fill=LoperamideTreatment))+scale_shape_manual(values = c(21,22,23,24,25))+ -->
<!--   annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") + -->
<!--   # add the global p-value from the adonis2 function -->
<!--   geom_text(data=adonissub1, aes(x=0.5,y=0.8,label=paste("p=",`Pr(>F)`[1],"*")), color="gray40")+ -->
<!--   scale_fill_manual(values=c("navy","aquamarine4"))+scale_colour_manual(values=c("navy","aquamarine4"))+ -->
<!--   theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right", -->
<!--         legend.title = element_blank(),legend.box="horizontal")+ggtitle("Timepoint 1. 24h") -->
<!-- loperamide1sub -->


<!-- ### -->

<!-- metadataLoperamide2sub <- metadataLoperamide2  %>% filter(Loperamide!="Control water") -->
<!-- SVsLoperamidesub2<- matrixLoperamidePercent[c(21:30),] -->
<!-- adonissub2<-adonis2(SVsLoperamidesub2~Loperamide, data=metadataLoperamide2sub, by=NULL, method="bray", k=2) -->

<!-- sol<-metaMDS(SVsLoperamidesub2,distance = "bray", k = 2, trymax = 200) -->
<!-- sol$stress; stressplot(sol) -->

<!-- NMDS=data.frame(x=sol$point[,1],y=sol$point[,2], -->
<!--                 LoperamideTreatment=as.factor(metadataLoperamide2sub$Loperamide), -->
<!--                 TimepointDay=as.factor(metadataLoperamide2sub$TimepointDay), -->
<!--                 Rep=as.factor(metadataLoperamide2sub$SampleRep)) -->
<!-- plot.new() -->
<!-- ord<-ordiellipse(sol, NMDS$LoperamideTreatment, display = "sites", kind ="sd", conf = 0.95, label = T) -->
<!-- dev.off() -->
<!-- df_ell <- data.frame() -->
<!-- for(g in levels(NMDS$LoperamideTreatment)){ -->
<!--   if(g!="" && (g %in% names(ord))){ -->
<!--     df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTreatment==g,], -->
<!--                                                      veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTreatment=g))}} -->
<!-- head(df_ell) -->
<!-- NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTreatment),mean) -->
<!-- loperamide2sub <-ggplot(data=NMDS,aes(x,y))+ -->
<!--   geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTreatment,colour=LoperamideTreatment), size=1) + -->
<!--   geom_point(size=4, alpha=0.9,aes(shape=LoperamideTreatment,fill=LoperamideTreatment))+scale_shape_manual(values = c(21,22,23,24,25))+ -->
<!--   annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") + -->
<!--   # add the global p-value from the adonis2 function -->
<!--   geom_text(data=adonissub2, aes(x=0.8,y=1,label=paste("p=",`Pr(>F)`[1],"*")), color="gray40")+ -->
<!--   scale_fill_manual(values=c("navy","aquamarine4"))+scale_colour_manual(values=c("navy","aquamarine4"))+ -->
<!--   theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right", -->
<!--         legend.title = element_blank(),legend.box="horizontal")+ggtitle("Timepoint 2. 48h") -->


<!-- ### -->

<!-- metadataLoperamide3sub <- metadataLoperamide3  %>% filter(Loperamide!="Control water") -->
<!-- SVsLoperamidesub3 <- matrixLoperamidePercent[c(36:45),] -->
<!-- adonissub3<-adonis2(SVsLoperamidesub3~Loperamide, data=metadataLoperamide3sub, by=NULL, method="bray", k=2) -->

<!-- sol<-metaMDS(SVsLoperamidesub3,distance = "bray", k = 2, trymax = 200) -->
<!-- sol$stress; stressplot(sol) -->

<!-- NMDS=data.frame(x=sol$point[,1],y=sol$point[,2], -->
<!--                 LoperamideTreatment=as.factor(metadataLoperamide3sub$Loperamide), -->
<!--                 TimepointDay=as.factor(metadataLoperamide3sub$TimepointDay), -->
<!--                 Rep=as.factor(metadataLoperamide3sub$SampleRep)) -->
<!-- plot.new() -->
<!-- ord<-ordiellipse(sol, NMDS$LoperamideTreatment, display = "sites", kind ="sd", conf = 0.95, label = T) -->
<!-- dev.off() -->
<!-- df_ell <- data.frame() -->
<!-- for(g in levels(NMDS$LoperamideTreatment)){ -->
<!--   if(g!="" && (g %in% names(ord))){ -->
<!--     df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTreatment==g,], -->
<!--                                                      veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTreatment=g))}} -->
<!-- head(df_ell) -->
<!-- NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTreatment),mean) -->
<!-- loperamide3sub <- ggplot(data=NMDS,aes(x,y))+ -->
<!--   geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTreatment,colour=LoperamideTreatment), size=1) + -->
<!--   geom_point(size=4, alpha=0.9,aes(shape=LoperamideTreatment,fill=LoperamideTreatment))+scale_shape_manual(values = c(21,22,23,24,25))+ -->
<!--   annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") + -->
<!--   # add the global p-value from the adonis2 function -->
<!--   geom_text(data=adonissub3, aes(x=1,y=0.75,label=paste("p=",`Pr(>F)`[1])), color="gray40")+ -->
<!--   scale_fill_manual(values=c("navy","aquamarine4"))+scale_colour_manual(values=c("navy","aquamarine4"))+ -->
<!--   theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right", -->
<!--         legend.title = element_blank(),legend.box="horizontal")+ggtitle("Timepoint 3. 6d") -->
<!-- ``` -->

<!-- ```{r plotbeta, fig.width=10, fig.height=4} -->
<!-- loperamide1sub+loperamide2sub+loperamide3sub + plot_layout(guides = "collect") &  -->
<!--    theme(legend.position = "bottom") #& coord_fixed() -->

<!-- loperamide1sub+loperamide2sub+loperamide3sub + plot_layout(guides = "collect") &  -->
<!--    theme(legend.position = "bottom", text=element_text(family="Avenir"), -->
<!--          panel.background = element_rect(size=1), -->
<!--          plot.background = element_blank(), legend.background = element_blank()) -->
<!-- ggsave("LoperamideBetaAllSeparateTimeline_avenir.png", bg="transparent", width = 12, height = 5) -->



<!-- ``` -->



# Beta Diversity

```{r betadivsetup}

# ellipse function
veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100)
{
   theta <- (0:npoints) * 2 * pi/npoints
   Circle <- cbind(cos(theta), sin(theta))
   t(center + scale * t(Circle %*% chol(cov)))
}

```

## All conditions

```{r loperamidebetaalltogether, fig.show='hide', results=FALSE, echo=FALSE}

set.seed(123)

metadataLoperamide <- filter(metadata, Project=="Loperamide")

sol<-metaMDS(matrixLoperamidePercent,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                LoperamideTreatment=as.factor(metadataLoperamide$Loperamide),
                TimepointDay=as.factor(metadataLoperamide$TimepointDay),
                Rep=as.factor(metadataLoperamide$SampleRep),
                LoperamideTimepoint=as.factor(metadataLoperamide$LoperamideTimepoint))
plot.new()
ord<-ordiellipse(sol, NMDS$LoperamideTimepoint, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$LoperamideTimepoint)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTimepoint==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTimepoint=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTimepoint),mean)
```

```{r betaplotallconditions1, fig.width=8}
ggplot(data=NMDS,aes(x,y,colour=LoperamideTimepoint,fill=LoperamideTimepoint))+theme_bw() +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTimepoint), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=TimepointDay))+scale_shape_manual(values = c(21,22,23,24,25))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") +
  scale_color_manual(values=c("grey80","grey50","grey20",'#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759')) +
  scale_fill_manual(values=c("grey80","grey50","grey20",'#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759')) +
  scale_linetype_manual(values=c("solid","solid","solid","dotted","dotted","dotted","dashed","dashed","dashed"))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")

df_ell <- df_ell %>% left_join(distinct(metadataLoperamide, 
                                        LoperamideTimepoint,TimepointDay, Loperamide))
NMDS <- NMDS %>% left_join(distinct(metadataLoperamide, 
                                        LoperamideTimepoint,TimepointDay, Loperamide))

allbetaplot <- ggplot(data=NMDS,aes(x,y))+
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, colour=Loperamide, lty=TimepointDay), size=1) +
  geom_point(aes(color=Loperamide, shape=TimepointDay),size=4, alpha=0.9)+
  scale_shape_manual(values=c(15,16,17))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5,
           color=c("#000000","#000000","#000000",'#1c5580','#1c5580',
                   '#1c5580','#0fc08e','#0fc08e','#0fc08e')) +
  annotate("richtext", label=paste0("*k* = 2<br>stress = ", round(sol$stress,3)), 
            hjust=0, x=-1.8, y=-1.3, color="grey20", label.size = NA, size = 6, fill=NA) + 
  scale_color_manual(values=c("#000000",'#1c5580','#0fc08e')) +
  scale_linetype_manual(values=c("solid","dotted","dashed","solid","dotted","dashed","solid","dotted","dashed"))+
  theme(legend.text = element_text(size=14, colour="gray20"), 
        legend.position = "bottom", legend.direction="vertical",
        legend.title = element_blank(),legend.box="horizontal",
        legend.key.width = unit(3,"cm"))+
   guides(color=guide_legend(override.aes = list(size=5, shape=15)))
allbetaplot
ggsave("LoperamideBetaAllWaterToo.png", bg="transparent", width = 10, height = 6)

```


```{r betaallpresent}
ggplot(data=NMDS,aes(x,y,colour=LoperamideTimepoint,fill=LoperamideTimepoint))+theme_bw() +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTimepoint), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=TimepointDay))+scale_shape_manual(values = c(21,22,23,24,25))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5,  family="Avenir",
           color=c("#000000","#000000","#000000",'#1c5580', '#1c5580', '#1c5580', 
                             '#0fc08e','#0fc08e','#0fc08e')) +
  scale_color_manual(values=c("#000000","#000000","#000000",'#1c5580', '#1c5580', '#1c5580', 
                             '#0fc08e','#0fc08e','#0fc08e')) +
  scale_fill_manual(values=c("#000000","#000000","#000000",'#1c5580', '#1c5580', '#1c5580', 
                             '#0fc08e','#0fc08e','#0fc08e')) +
  scale_linetype_manual(values=c("solid","solid","solid","dotted","dotted","dotted","dashed","dashed","dashed"))+
  theme(text = element_text(family="Avenir"),plot.background = element_blank(),
        legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")

#ggsave("LoperamideBetaAllWaterToo_avenir.png", bg="transparent", width = 10, height = 6)

```

## Just DMSO and loperamide


```{r loperamidedmsobeta, fig.show='hide', results=FALSE, echo=FALSE}

set.seed(123)

# subset metadata
metadataLoperamidesub <- filter(metadataLoperamide, Project=="Loperamide") %>%
      filter(Loperamide != "Control water")
# subset taxa matrix
SVsLoperamidesub<- matrixLoperamidePercent[c(6:15,21:30,36:45),]

sol<-metaMDS(SVsLoperamidesub,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

adonis2(SVsLoperamidesub~Loperamide, data=metadataLoperamidesub, by=NULL, method="bray", k=2)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                LoperamideTreatment=as.factor(metadataLoperamidesub$Loperamide),
                TimepointDay=as.factor(metadataLoperamidesub$TimepointDay),
                Rep=as.factor(metadataLoperamidesub$SampleRep),
                LoperamideTimepoint=as.factor(metadataLoperamidesub$LoperamideTimepoint))

plot.new()
ord<-ordiellipse(sol, NMDS$LoperamideTimepoint, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$LoperamideTimepoint)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTimepoint==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTimepoint=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTimepoint),mean)

```

```{r betaplotnowater, fig.width=6.5, fig.height=6}

betadivplot<- ggplot(data=NMDS,aes(x,y,colour=LoperamideTimepoint,fill=LoperamideTimepoint))+
   geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTimepoint), size=0.8) +
   geom_point(size=3, alpha=0.7, aes(shape=LoperamideTimepoint), color="grey30")+
   scale_shape_manual(values=c(24,24,24,22,22,22))+
   annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5,
           color=c('#5a89b8',  '#1c5580', '#00264c', '#0fc08e','#008759', '#008759')) +
   annotate("richtext", label=paste0("stress = ", round(sol$stress,3),"<br>*k* = 2"), 
            hjust=0, x=1, y=-1, color="grey20", label.size = NA, size = 6, fill=NA) + 
   scale_color_manual(values=c('#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759')) +
   scale_fill_manual(values=c('#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759')) +
   scale_linetype_manual(values=c("solid","dotted","dashed","solid","dotted","dashed"))+
   theme(legend.text = element_text(size=18), legend.position="bottom",
        legend.title = element_blank(),legend.direction="vertical", legend.key.width = unit(2,"cm"))+ 
   guides(shape = guide_legend(ncol = 2))
betadivplot
ggsave("LoperamideBetaAll.png", bg="transparent", width = 6.5, height = 8)
ggsave("LoperamideBetaAll.svg", bg="transparent", width = 8, height = 4)

# for presentation
ggplot(data=NMDS,aes(x,y,colour=LoperamideTimepoint,fill=LoperamideTimepoint))+
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTimepoint), size=1.5) +
  geom_point(size=4, alpha=0.9,aes(shape=LoperamideTimepoint))+
   scale_shape_manual(values=c(21,24,22,21,24,22))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, face="bold",
           color=c('#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759'), 
           family="Avenir") +
   scale_color_manual(values=c('#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759')) +
   scale_fill_manual(values=c('#5a89b8',  '#1c5580', '#00264c', '#69fcc6','#0fc08e', '#008759')) +
   scale_linetype_manual(values=c("solid","dotted","dashed","solid","dotted","dashed"))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "bottom", 
        text=element_text(family="Avenir"),
        legend.title = element_blank(),legend.box="vertical",legend.key.width = unit(2,"cm"))+ 
   guides(shape = guide_legend(ncol = 2))

#ggsave("LoperamideBetaAll_avenir.png", bg="transparent", width = 8, height = 8)

```

#### With limma results

```{r betadivlimma, fig.height=10, fig.width=14}


betadivplot+limmagenus+
   plot_annotation(tag_levels = "A",
                   theme = theme(plot.title = element_text(size = 20)))+
   plot_layout(widths=c(4,3))&
   theme(plot.tag = element_text(face = "bold", size=20))
ggsave("LoperamideBetaAll_LimmaGenus.png", bg="transparent", width = 15, height = 10)


```





## By treatment and then timepoint

```{r loperamidebetaall, fig.show='hide', results=FALSE, echo=FALSE}

sol<-metaMDS(matrixLoperamidePercent,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                Loperamide=as.factor(metadataLoperamide$Loperamide),
                TimepointDay=as.factor(metadataLoperamide$TimepointDay),
                Rep=as.factor(metadataLoperamide$SampleRep))
plot.new()
ord<-ordiellipse(sol, NMDS$Loperamide, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$Loperamide)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$Loperamide==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),Loperamide=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$Loperamide),mean)
loperamidealltreat <- ggplot(data=NMDS,aes(x,y,colour=Loperamide,fill=Loperamide))+theme_bw() +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=Loperamide), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=TimepointDay))+   scale_shape_manual(values=c(21,24,22))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") +
  scale_fill_manual(values=c('#000000','#1c5580', '#0fc08e'))+
   scale_colour_manual(values=c('#000000','#1c5580', '#0fc08e'))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")

###

plot.new()
ord<-ordiellipse(sol, NMDS$TimepointDay, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$TimepointDay)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$TimepointDay==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),TimepointDay=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$TimepointDay),mean)
loperamidealltime <- ggplot(data=NMDS,aes(x,y,colour=TimepointDay,fill=TimepointDay))+theme_bw() +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=TimepointDay), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=Loperamide))+scale_shape_manual(values = c(21,22,23,24,25))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") +
  scale_fill_manual(values=c("rosybrown1","firebrick2","firebrick4"))+scale_colour_manual(values=c("rosybrown1","firebrick2","firebrick4"))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")
```

```{r betaplotbytreatment}
library(patchwork)
loperamidealltreat+loperamidealltime  & theme(legend.position = "bottom", legend.direction = "vertical")

```

## Each timepoint separately


```{r loperamidebeta, fig.show='hide', results=FALSE, echo=FALSE}

### T0

metadataLoperamide1 <- filter(metadata, Project=="Loperamide") %>% filter(Timepoint==1)
SVsLoperamidesub1<- matrixLoperamidePercent[c(1:15),]

sol<-metaMDS(SVsLoperamidesub1,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                LoperamideTreatment=as.factor(metadataLoperamide1$Loperamide),
                TimepointDay=as.factor(metadataLoperamide1$TimepointDay),
                Rep=as.factor(metadataLoperamide1$SampleRep))
plot.new()
ord<-ordiellipse(sol, NMDS$LoperamideTreatment, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$LoperamideTreatment)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTreatment==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTreatment=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTreatment),mean)
loperamide1 <- ggplot(data=NMDS,aes(x,y,colour=LoperamideTreatment,fill=LoperamideTreatment))+theme_bw() +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTreatment), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=LoperamideTreatment), color="black")+
   scale_shape_manual(values=c(21,24,22))+
   annotate("richtext", label=paste0("stress = ", round(sol$stress,3)), 
            hjust=0, x=-1.7, y=-0.8, color="gray40", label.size = NA, size = 4, fill=NA) + 
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") +
  scale_fill_manual(values=c('#000000','#1c5580', '#0fc08e'))+
   scale_colour_manual(values=c('#000000','#1c5580', '#0fc08e'))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")+ggtitle("T0")

### T1

metadataLoperamide2 <- filter(metadata, Project=="Loperamide") %>% filter(Timepoint==2)
SVsLoperamidesub2<- matrixLoperamidePercent[c(16:30),]

sol<-metaMDS(SVsLoperamidesub2,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                LoperamideTreatment=as.factor(metadataLoperamide2$Loperamide),
                TimepointDay=as.factor(metadataLoperamide2$TimepointDay),
                Rep=as.factor(metadataLoperamide2$SampleRep))
plot.new()
ord<-ordiellipse(sol, NMDS$LoperamideTreatment, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$LoperamideTreatment)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTreatment==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTreatment=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTreatment),mean)
loperamide2 <-ggplot(data=NMDS,aes(x,y,colour=LoperamideTreatment,fill=LoperamideTreatment))+theme_bw() +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTreatment), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=LoperamideTreatment), color="black")+
   scale_shape_manual(values=c(21,24,22))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") +
   annotate("richtext", label=paste0("stress = ", round(sol$stress,3)), 
            hjust=0, x=0.4, y=-1.2, color="gray40", label.size = NA, size = 4, fill=NA) + 
  scale_fill_manual(values=c('#000000','#1c5580', '#0fc08e'))+
   scale_colour_manual(values=c('#000000','#1c5580', '#0fc08e'))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")+ggtitle("T1")


### T5


metadataLoperamide3 <- filter(metadata, Project=="Loperamide") %>% filter(Timepoint==3)
SVsLoperamidesub3<- matrixLoperamidePercent[c(31:45),]

sol<-metaMDS(SVsLoperamidesub3,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                LoperamideTreatment=as.factor(metadataLoperamide3$Loperamide),
                TimepointDay=as.factor(metadataLoperamide3$TimepointDay),
                Rep=as.factor(metadataLoperamide3$SampleRep))
plot.new()
ord<-ordiellipse(sol, NMDS$LoperamideTreatment, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$LoperamideTreatment)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTreatment==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTreatment=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTreatment),mean)
loperamide3 <- ggplot(data=NMDS,aes(x,y,colour=LoperamideTreatment,fill=LoperamideTreatment))+theme_bw() +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTreatment), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=LoperamideTreatment), color="black")+
   scale_shape_manual(values=c(21,24,22))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size = 5, color="gray40") +
   annotate("richtext", label=paste0("stress = ", round(sol$stress,3)), 
            hjust=0, x=-1.2, y=-0.66, color="gray40", label.size = NA, size = 4, fill=NA) + 
  scale_fill_manual(values=c('#000000','#1c5580', '#0fc08e'))+
   scale_colour_manual(values=c('#000000','#1c5580', '#0fc08e'))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")+ggtitle("T5")
```

```{r betaallplot1,  fig.width=12, fig.height=6}

library(patchwork)
loperamide1+loperamide2+loperamide3 + plot_layout(guides = "collect") & 
   theme(legend.position = "bottom", text=element_text(family="Avenir"),
         plot.background = element_blank(), legend.background = element_blank()) & coord_fixed()
#ggsave("LoperamideBetaAllSeparate_avenir.png", bg="transparent", width = 12, height = 6)
loperamide1 +loperamide2 + loperamide3 + guide_area() + plot_layout(guides = "collect") & 
   theme(legend.position = "right", text=element_text(family="Avenir"),
         plot.background = element_blank(), legend.background = element_blank())
#ggsave("LoperamideBetaAllSeparate3_avenir.png", bg="transparent", width = 7, height = 6)


betaindiv<- loperamide1 +loperamide2 + loperamide3 + plot_layout(guides = "collect") & 
   theme(legend.position = "right",legend.key.width = unit(2,"cm"),
         plot.title=element_text(size=24),
         plot.background = element_blank(), legend.background = element_blank())
ggsave("LoperamideBetaAllSeparate3.png", bg="transparent", width = 7, height = 6)

```

## Each timepoint separately, just DMSO and loperamide


```{r loperamidebetadmsolop,fig.show='hide', results=FALSE, echo=FALSE}

metadataLoperamide1sub <- metadataLoperamide1  %>% filter(Loperamide!="Control water")
SVsLoperamidesub1<- matrixLoperamidePercent[c(6:15),]
adonissub1<-adonis2(SVsLoperamidesub1~Loperamide, data=metadataLoperamide1sub, by=NULL, method="bray", k=2)

sol<-metaMDS(SVsLoperamidesub1,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                LoperamideTreatment=as.factor(metadataLoperamide1sub$Loperamide),
                TimepointDay=as.factor(metadataLoperamide1sub$TimepointDay),
                Rep=as.factor(metadataLoperamide1sub$SampleRep))
plot.new()
ord<-ordiellipse(sol, NMDS$LoperamideTreatment, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$LoperamideTreatment)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTreatment==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTreatment=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTreatment),mean)
loperamide1sub <- ggplot(data=NMDS,aes(x,y))+
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTreatment,colour=LoperamideTreatment), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=LoperamideTreatment,fill=LoperamideTreatment))+scale_shape_manual(values = c(21,22,23,24,25))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") +
  # add the global p-value from the adonis2 function
  geom_text(data=adonissub1, aes(x=0.5,y=0.8,label=paste("p=",`Pr(>F)`[1],"*")), color="gray40")+
  scale_fill_manual(values=c('#1c5580', '#0fc08e'))+scale_colour_manual(values=c('#1c5580', '#0fc08e'))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")+ggtitle("Timepoint 1. 24h")
loperamide1sub


###

metadataLoperamide2sub <- metadataLoperamide2  %>% filter(Loperamide!="Control water")
SVsLoperamidesub2<- matrixLoperamidePercent[c(21:30),]
adonissub2<-adonis2(SVsLoperamidesub2~Loperamide, data=metadataLoperamide2sub, by=NULL, method="bray", k=2)

sol<-metaMDS(SVsLoperamidesub2,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                LoperamideTreatment=as.factor(metadataLoperamide2sub$Loperamide),
                TimepointDay=as.factor(metadataLoperamide2sub$TimepointDay),
                Rep=as.factor(metadataLoperamide2sub$SampleRep))
plot.new()
ord<-ordiellipse(sol, NMDS$LoperamideTreatment, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$LoperamideTreatment)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTreatment==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTreatment=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTreatment),mean)
loperamide2sub <-ggplot(data=NMDS,aes(x,y))+
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTreatment,colour=LoperamideTreatment), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=LoperamideTreatment,fill=LoperamideTreatment))+scale_shape_manual(values = c(21,22,23,24,25))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") +
  # add the global p-value from the adonis2 function
  geom_text(data=adonissub2, aes(x=0.8,y=1,label=paste("p=",`Pr(>F)`[1],"*")), color="gray40")+
  scale_fill_manual(values=c('#1c5580', '#0fc08e'))+scale_colour_manual(values=c('#1c5580', '#0fc08e'))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")+ggtitle("Timepoint 2. 48h")


###

metadataLoperamide3sub <- metadataLoperamide3  %>% filter(Loperamide!="Control water")
SVsLoperamidesub3 <- matrixLoperamidePercent[c(36:45),]
adonissub3<-adonis2(SVsLoperamidesub3~Loperamide, data=metadataLoperamide3sub, by=NULL, method="bray", k=2)

sol<-metaMDS(SVsLoperamidesub3,distance = "bray", k = 2, trymax = 200)
sol$stress; stressplot(sol)

NMDS=data.frame(x=sol$point[,1],y=sol$point[,2],
                LoperamideTreatment=as.factor(metadataLoperamide3sub$Loperamide),
                TimepointDay=as.factor(metadataLoperamide3sub$TimepointDay),
                Rep=as.factor(metadataLoperamide3sub$SampleRep))
plot.new()
ord<-ordiellipse(sol, NMDS$LoperamideTreatment, display = "sites", kind ="sd", conf = 0.95, label = T)
dev.off()
df_ell <- data.frame()
for(g in levels(NMDS$LoperamideTreatment)){
  if(g!="" && (g %in% names(ord))){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$LoperamideTreatment==g,],
                                                     veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale))),LoperamideTreatment=g))}}
head(df_ell)
NMDS.mean=aggregate(NMDS[,1:2],list(group=NMDS$LoperamideTreatment),mean)
loperamide3sub <- ggplot(data=NMDS,aes(x,y))+
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2, lty=LoperamideTreatment,colour=LoperamideTreatment), size=1) +
  geom_point(size=4, alpha=0.9,aes(shape=LoperamideTreatment,fill=LoperamideTreatment))+scale_shape_manual(values = c(21,22,23,24,25))+
  annotate("text",x=NMDS.mean$x,y=NMDS.mean$y,label=NMDS.mean$group,size=5, color="gray40") +
  # add the global p-value from the adonis2 function
  geom_text(data=adonissub3, aes(x=1,y=0.75,label=paste("p=",`Pr(>F)`[1])), color="gray40")+
  scale_fill_manual(values=c('#1c5580', '#0fc08e'))+scale_colour_manual(values=c('#1c5580', '#0fc08e'))+
  theme(legend.text = element_text(size=14, colour="gray20"), legend.position = "right",
        legend.title = element_blank(),legend.box="horizontal")+ggtitle("Timepoint 3. 6d")
```

```{r plotbeta, fig.width=10, fig.height=4}
loperamide1sub+loperamide2sub+loperamide3sub + plot_layout(guides = "collect") & 
   theme(legend.position = "bottom") #& coord_fixed()

loperamide1sub+loperamide2sub+loperamide3sub + plot_layout(guides = "collect") & 
   theme(legend.position = "bottom", text=element_text(family="Avenir"),
         panel.background = element_rect(size=1),
         plot.background = element_blank(), legend.background = element_blank())
ggsave("LoperamideBetaAllSeparateTimeline_avenir.png", bg="transparent", width = 12, height = 5)



```


## Within group beta-diversity

```{r withinbeta}

braycurtisdistances<-
  # calculate the dissimilarity matrix between each sample
  vegdist(matrixLoperamidePercent, method="bray", k=2) %>% as.matrix() %>% as.data.frame() %>% 
  # Add in the sites based on the rows (paired sample)
  rownames_to_column(var="PairedID")  %>% mutate(PairedCondition=metadataLoperamide$LoperamideTimepoint) %>% 
  # Make into longform based on the columns (OG SampleID)
  pivot_longer(cols="1-Conv-1-1":"45-LOP-3-5", names_to="SampleID") %>% 
  # Add in sites based on the columns (OG SampleID)
  left_join(metadataLoperamide) %>% 
  # Remove rows where a sample is paired with a sample from another condition
  filter(PairedCondition==LoperamideTimepoint) %>% 
  # Remove the diagonal rows, where each sample was compared to itself
  filter(value!=0)

# stats
compare_means(data=braycurtisdistances, value~LoperamideTimepoint, method="wilcox", p.adjust.method = "BH")
compare_means(data=braycurtisdistances, value~LoperamideTimepoint, method="kruskal")

# plot within condition variation
withindiss <- ggplot(braycurtisdistances, aes(x=Loperamide, y=value, fill=Loperamide, shape=Loperamide))+
   geom_jitter(width=0.15, size=2, alpha=0.7)+geom_boxplot(alpha=0.8)+
   labs(x=NULL,y="within condition dissimilarity index",fill="Condition")+
   facet_grid(.~TimepointDay, space="free", scales="free")+
   stat_compare_means(ref.group="DMSO", label="p.signif")+
   scale_fill_manual(values=c('#000000','#1c5580', '#0fc08e'))+
   scale_shape_manual(values=c(21,24,22))+
   scale_y_continuous(limits=c(0,0.8), labels=c("0.00","0.25","0.50","0.75","1.00"))+
   theme(legend.position = "none", strip.background = element_rect("grey85", color="transparent"))

# for presentation
ggplot(braycurtisdistances, aes(x=Loperamide, y=value, fill=Loperamide))+
  geom_jitter(width=0.15, size=2, shape=23, alpha=0.7)+geom_boxplot(alpha=0.8)+
  labs(x=NULL,y="within condition dissimilarity index",fill="Condition")+
   facet_grid(.~TimepointDay, space="free", scales="free")+
   stat_compare_means(ref.group="DMSO", label="p.signif")+
   scale_shape_manual(values=c(21,24,22))+
  scale_fill_manual(values=c('#000000','#1c5580', '#0fc08e'))+
  scale_y_continuous(limits=c(0,1.19), labels=c("0.00","0.25","0.50","0.75","1.00"," "))+
   theme(text=element_text(family="Avenir", size=14), axis.text.x = element_blank(),
         axis.ticks.x = element_blank(), 
         legend.position = "right", strip.background = element_blank(),
         panel.background = element_rect(color="transparent"),
         legend.text = element_text(size=16))

ggsave("16Sloperamide_withinBray_avenir.png", bg="transparent", width=10, height=5)
```


#### Beta diversity summary plot

```{r betasummary, fig.width=15, fig.height=25}
(allbetaplot  & theme(legend.position=c(0.25,0.86))) /
   (betaindiv  & theme(legend.position="bottom"))/
   (withindiss & theme()) +plot_annotation(tag_levels = c('A','B'))+
   plot_layout(heights=c(3.5,2,2.5)) &
   theme(plot.tag = element_text(face = "bold", size=20))
ggsave("16Sloperamide_BetaDaysAndWithinDiv.png", bg="transparent", width=11, height=16)



```


# Session Info

```{r session}

sessionInfo()


```


