---
title: "Water Loperamide CFUs"
author: "Rebecca Stevick"
date: "updated 2022.04.04"
output: 
   html_document:
      toc: TRUE
      theme: "cerulean"
      toc_float:
         collapsed: false
         smooth_scroll: false

---

## About the Data

Individual strains survival in water, in vitro assay performed from 20220301 - 20220307. Second round from 20220308-0314. Third round from 20220321-0324. Fourth round 20220516-0520.

20 mL of volvic water with each strain at final OD=0.001 (~10^6^ CFU/mL)\
DMSO at 1:10000, Loperamide at 10 mg/L

### Timepoints

Add loperamide and strains to water.
Sample water at 0 h, 3 h, 6 hours on day 1.
Then, daily until 3 days post inoculation. 24 h, 48 h, 72 h.


### Sample collection & plating

At each timepoint:

Make 0 to -3 dilutions in 96-well plates, in triplicate\
Plate 10 µL microdrops on big square plates.\
2 square plates total per timepoint


Put plates at 28C for 2 days, then count colonies.

### Conditions

**Treatment (Strain)**: Bc1, Bc2, Bc3, Bc4, Bc10, W6, W8, Mz8, Fjohn

**Loperamide Treatment**: Control (water only), DMSO, Loperamide 10 mg/L



## Setup

### Load libraries and settings

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(scales)
library(ggpubr)
library(ungeviz)
library(ggtext)
library(rmdformats)

# set global theme
theme_set(theme_minimal()+
theme(panel.grid.major.y = element_line(color="grey80"), strip.text=element_text(size=16),
        strip.text.y = element_text(angle=0), plot.caption = element_text(size=10),
        panel.grid.major.x = element_blank(),legend.position="top",
        plot.background = element_rect(fill="transparent", color="transparent"),
        axis.ticks = element_line(inherit.blank = FALSE),
        panel.background = element_rect(color="grey50", size=2), 
        legend.title = element_text(size=18),
        axis.text = element_text(size=15), axis.title = element_text(size=18),
        legend.text = element_text(size=16), plot.title = element_text(hjust=0.5)))

# set global options for code output
knitr::opts_chunk$set(echo=TRUE, warning=FALSE,message=FALSE)

```

### Import data

```{r import, echo=TRUE}

# import individual strain data from exp 1
datacfus1 <-
   readxl::read_xlsx("LoperamideWaterSurvivalCFUs.xlsx", sheet="Round1") %>%
   drop_na(DF) %>%
   mutate(LoperamideTreatment=factor(LoperamideTreatment, 
                                    levels=c("None", "DMSO", "Loperamide 10 mg/L"),
                                    labels=c("Control","DMSO", "Loperamide")),
         Treatment = factor(Treatment, levels=c("Bc1","Bc2","Bc3","Bc4","Bc10"))) %>% 
   add_column(Assay=1)

# import individual strain data from exp 2
datacfus2 <-
   readxl::read_xlsx("LoperamideWaterSurvivalCFUs.xlsx", sheet="Round2") %>%
   drop_na(DF) %>%
   mutate(LoperamideTreatment=factor(LoperamideTreatment, 
                                    levels=c("None", "DMSO", "Loperamide 10 mg/L"),
                                    labels=c("Control","DMSO", "Loperamide")),
         Treatment = factor(Treatment, levels=c("Bc1","Bc2","Bc3","Bc4","Bc10"))) %>% 
   add_column(Assay=2)

# import individual strain data from exp 3
datacfus3 <-
   readxl::read_xlsx("LoperamideWaterSurvivalCFUs.xlsx", sheet="Round3") %>%
   drop_na(DF) %>%
   mutate(LoperamideTreatment=factor(LoperamideTreatment, 
                                    levels=c("None", "DMSO", "Loperamide 10 mg/L"),
                                    labels=c("Control","DMSO", "Loperamide")),
         Treatment = factor(Treatment, levels=c("Bc1","Bc2","Bc3","Bc4","Bc10"))) %>% 
   add_column(Assay=3)


# import individual strain data from exp 4
datacfus4 <-
   readxl::read_xlsx("LoperamideWaterSurvivalCFUs.xlsx", sheet="Round4") %>%
   drop_na(DF) %>%
   mutate(LoperamideTreatment=factor(LoperamideTreatment, 
                                    levels=c("None", "DMSO", "Loperamide 10 mg/L"),
                                    labels=c("Control","DMSO", "Loperamide")),
         Treatment = factor(Treatment, levels=c("W6","W8","Mz8")))

datacfus5 <-
   readxl::read_xlsx("LoperamideWaterSurvivalCFUs.xlsx", sheet="Round5") %>%
   drop_na(DF) %>%
   mutate(LoperamideTreatment=factor(LoperamideTreatment, 
                                    levels=c("None", "DMSO", "Loperamide 10 mg/L"),
                                    labels=c("Control","DMSO", "Loperamide")),
         Treatment = factor(Treatment, levels=c("W6","W8","Mz8","Fjohn","Mz1")))


straininfo <- readxl::read_xlsx("../LoperamideStrainInfo.xlsx") %>% 
   mutate(Strain=recode(Strain, "W6t"="W6"))


dataallCFUs <- full_join(datacfus1, datacfus2 %>% filter(Treatment!="Bc4")) %>% 
   full_join(datacfus3) %>% 
   full_join(datacfus4) %>% full_join(datacfus5) %>% 
   left_join(straininfo, by=c("Treatment"="Strain")) %>% 
   mutate(CodeName=factor(CodeName, levels=unique(straininfo$CodeName)))


```

------------------------------------------------------------------------


## Stats of all significant comparisons

```{r strainstats}

longdata <- dataallCFUs %>% 
  pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  select(Treatment, LoperamideTreatment, Timepoint_hrs, RepCFUs)

statsbydayCFUs <- compare_means(data=longdata, 
                            RepCFUs~LoperamideTreatment, 
                            group.by = c("Treatment","Timepoint_hrs")) %>% 
   left_join(straininfo, by=c("Treatment"="Strain")) %>% 
   mutate(CodeName=factor(CodeName, levels=unique(straininfo$CodeName)))
gt::gt(statsbydayCFUs %>% filter(p.format<0.05 & group1=="DMSO"))

```


## Experiment 1

### Timeline for control conditions

```{r timelineplotcontrol, fig.height=5, fig.width=7}
datacfus1 %>% filter(LoperamideTreatment=="Control") %>% 
  pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x=Timepoint_hrs, y=RepCFUs, 
             color=Treatment, fill=Treatment, shape=Treatment))+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=0.5, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=Treatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', limits=c(1e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed", "dotdash","longdash"))+
  scale_shape_manual(values=c(21,22,23,24,25))+
  scale_fill_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  scale_color_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  labs(y="CFUs per mL", x="Time (hours)", color=NULL, shape=NULL, fill=NULL, lty=NULL,
       title="Survival in water, control conditions")+
  theme(legend.key.width = unit(1.33,"cm"), legend.position = "right")
#ggsave("WaterInVitro_LoperamideCFUs_control1.png", bg="transparent", width = 7, height = 5)
```

### Timeline for each strain with mean

```{r timelineplot, fig.height=5, fig.width=11}
datacfus1 %>% pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = Timepoint_hrs, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_grid(.~Treatment)+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=1, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', limits=c(1e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c("grey20","navy","aquamarine4"))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", color=NULL, shape=NULL, lty=NULL)+
  theme(legend.key.width = unit(1.33,"cm"))
#ggsave("WaterInVitro_LoperamideCFUs1.png", bg="transparent", width = 12, height = 5)
```


### Mean CFUs only

```{r waterplot, fig.width=11, fig.height=5}
datacfus1 %>% 
  ggplot(aes(x = Timepoint_hrs, y=CFUs_permL, 
             lty=LoperamideTreatment, color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_grid(.~Treatment)+
  geom_point(size=3) +  geom_line(lwd=0.8) +
  scale_y_continuous(trans = 'log10', limits=c(1e4,1e7), # expand=c(0,0),
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_color_manual(values=c("grey20","navy","aquamarine4"))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", color=NULL, shape=NULL, lty=NULL)+
  theme(legend.key.width = unit(1.33,"cm"))
#ggsave("WaterInVitro_LoperamideCFUs2.png", bg="transparent", width = 11, height = 6)
```





## Experiment 2

### Timeline for control conditions

```{r timelineplotcontrol2, fig.height=5, fig.width=7}
datacfus2 %>% filter(LoperamideTreatment=="Control") %>% 
  pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x=Timepoint_hrs, y=RepCFUs, 
             color=Treatment, fill=Treatment, shape=Treatment))+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=0.5, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=Treatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', #limits=c(1e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed", "dotdash","longdash"))+
  scale_shape_manual(values=c(21,22,23,24,25))+
  scale_fill_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  scale_color_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  labs(y="CFUs per mL", x="Time (hours)", color=NULL, shape=NULL, fill=NULL, lty=NULL,
       title="Survival in water, control conditions")+
  theme(legend.key.width = unit(1.33,"cm"), legend.position = "right")
#ggsave("WaterInVitro_LoperamideCFUs_control2.png", bg="transparent", width = 7, height = 5)
```

### Timeline for each strain with mean

```{r timelineplot2, fig.height=5, fig.width=11}
datacfus2 %>% pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = Timepoint_hrs, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_grid(.~Treatment)+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=1, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', limits=c(NA, 1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c("grey20","navy","aquamarine4"))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", color=NULL, shape=NULL, lty=NULL)+
  theme(legend.key.width = unit(1.33,"cm"))
#ggsave("WaterInVitro_LoperamideCFUs2.png", bg="transparent", width = 12, height = 5)
```



## Experiment 3

### Timeline for control conditions

```{r timelineplotcontrol3, fig.height=5, fig.width=7}
datacfus3 %>% filter(LoperamideTreatment=="Control") %>% 
  pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x=Timepoint_hrs, y=RepCFUs, 
             color=Treatment, fill=Treatment, shape=Treatment))+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=0.5, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=Treatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', #limits=c(1e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed", "dotdash","longdash"))+
  scale_shape_manual(values=c(21,22,23,24,25))+
  scale_fill_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  scale_color_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  labs(y="CFUs per mL", x="Time (hours)", color=NULL, shape=NULL, fill=NULL, lty=NULL,
       title="Survival in water, control conditions")+
  theme(legend.key.width = unit(1.33,"cm"), legend.position = "right")
#ggsave("WaterInVitro_LoperamideCFUs_control3.png", bg="transparent", width = 7, height = 5)
```

### Timeline for each strain with mean

```{r timelineplot3, fig.height=5, fig.width=11}
datacfus3 %>% pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = Timepoint_hrs, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_grid(.~Treatment)+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=1, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', limits=c(NA, 1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c("grey20","navy","aquamarine4"))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", color=NULL, shape=NULL, lty=NULL)+
  theme(legend.key.width = unit(1.33,"cm"))
#ggsave("WaterInVitro_LoperamideCFUs3.png", bg="transparent", width = 12, height = 5)
```


## Experiment 4

### Timeline for control conditions

```{r timelineplotcontrol4, fig.height=5, fig.width=7}
datacfus4 %>% filter(LoperamideTreatment=="Control") %>% 
  pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x=Timepoint_hrs, y=RepCFUs, 
             color=Treatment, fill=Treatment, shape=Treatment))+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=0.5, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=Treatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', #limits=c(1e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed", "dotdash","longdash"))+
  scale_shape_manual(values=c(21,22,23,24,25))+
  scale_fill_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  scale_color_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  labs(y="CFUs per mL", x="Time (hours)", color=NULL, shape=NULL, fill=NULL, lty=NULL,
       title="Survival in water, control conditions")+
  theme(legend.key.width = unit(1.33,"cm"), legend.position = "right")
#ggsave("WaterInVitro_LoperamideCFUs_control3.png", bg="transparent", width = 7, height = 5)
```

### Timeline for each strain with mean

```{r timelineplot4, fig.height=5, fig.width=11}
datacfus4 %>% pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = Timepoint_hrs, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_grid(.~Treatment)+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=1, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', limits=c(NA, 1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c("grey20","navy","aquamarine4"))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", color=NULL, shape=NULL, lty=NULL)+
  theme(legend.key.width = unit(1.33,"cm"))
#ggsave("WaterInVitro_LoperamideCFUs3.png", bg="transparent", width = 12, height = 5)
```


## Experiment 5

### Timeline for control conditions

```{r timelineplotcontrol5, fig.height=5, fig.width=7}
datacfus5 %>% filter(LoperamideTreatment=="Control") %>% 
  pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x=Timepoint_hrs, y=RepCFUs, 
             color=Treatment, fill=Treatment, shape=Treatment))+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=0.5, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=Treatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', #limits=c(1e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed", "dotdash","longdash"))+
  scale_shape_manual(values=c(21,22,23,24,25))+
  scale_fill_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  scale_color_manual(values=nationalparkcolors::park_palette("Saguaro", n=5))+
  labs(y="CFUs per mL", x="Time (hours)", color=NULL, shape=NULL, fill=NULL, lty=NULL,
       title="Survival in water, control conditions")+
  theme(legend.key.width = unit(1.33,"cm"), legend.position = "right")
#ggsave("WaterInVitro_LoperamideCFUs_control3.png", bg="transparent", width = 7, height = 5)
```

### Timeline for each strain with mean

```{r timelineplot5, fig.height=5, fig.width=11}
datacfus5 %>% pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = Timepoint_hrs, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_grid(.~Treatment)+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=1, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', limits=c(NA, 1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c("grey20","navy","aquamarine4"))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", color=NULL, shape=NULL, lty=NULL)+
  theme(legend.key.width = unit(1.33,"cm"))
#ggsave("WaterInVitro_LoperamideCFUs3.png", bg="transparent", width = 12, height = 5)
```




## All together now

### Timeline for control conditions

```{r timelineplotcontrolall, fig.height=5, fig.width=7}
dataallCFUs %>% filter(LoperamideTreatment=="Control") %>% filter(Timepoint_hrs<80) %>% 
  pivot_longer(Rep1:Rep3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
   drop_na(PaperCode) %>% 
  ggplot(aes(x=Timepoint_hrs, y=RepCFUs, 
             color=PaperCode, fill=PaperCode, shape=PaperCode))+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=0.5, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=PaperCode), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', #limits=c(1e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
 # scale_linetype_manual(values=c("solid","dotted","dashed", "dotdash","longdash"))+
  scale_shape_manual(values=c(21,22,23,24,25,21,22,23,24,25))+
  scale_fill_manual(values=c(nationalparkcolors::park_palette("Saguaro", n=6),"grey","navy","lightcoral","deepskyblue"))+
  scale_color_manual(values=c(nationalparkcolors::park_palette("Saguaro", n=6),"grey","navy","lightcoral","deepskyblue"))+
  labs(y="CFUs per mL", x="Time (hours)", color=NULL, shape=NULL, fill=NULL, lty=NULL,
       title="Survival in water, control conditions")+
  theme(legend.key.width = unit(1.33,"cm"), legend.position = "right")
#ggsave("WaterInVitro_LoperamideCFUs_control_all.png", bg="transparent", width = 7, height = 5)
```

### Timeline for each strain with mean

```{r timelineplotall, fig.height=7, fig.width=15}

statsbydayCFUsformat <- statsbydayCFUs %>% filter(p.signif!="ns")  %>% filter(Timepoint_hrs<80)  %>%
   filter(Timepoint_hrs!=0) 

dataallCFUs %>% pivot_longer(Rep1:Rep3) %>% filter(Timepoint_hrs<80) %>% 
   filter(Timepoint_hrs !=3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = Timepoint_hrs, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_wrap(.~CodeName, ncol=5)+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=1, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=4) +
  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=1) +
  geom_text(data=statsbydayCFUsformat, aes(label="*", y=5e6, color=NA, shape=NA), size=9, color="#0fc08e",
            show.legend=FALSE)+
  scale_y_continuous(trans = 'log10', limits=c(NA,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
   scale_x_continuous(breaks=c(0,6,24,48,72))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", 
     #  title="All 3 experiments together",
       color=NULL, shape=NULL, lty=NULL)+
  theme(strip.text = element_markdown(size = 16, face="bold"), 
        panel.grid.major.x = element_line(size=0.5),
        panel.grid.minor.x = element_blank(),panel.grid.minor.y = element_blank(),
        legend.text = element_text(size=18), legend.key.width = unit(2,"cm"))

ggsave("WaterInVitro_LoperamideCFUs.png", bg="transparent", width = 19, height = 9.5)

```

```{r timelineplotalllong, fig.height=15, fig.width=5}

statsbydayCFUsformat <- statsbydayCFUs %>% filter(Timepoint_hrs %in% c(6,24)) %>% 
   filter(p.signif!="ns") 

survivalsublong<-dataallCFUs %>% pivot_longer(Rep1:Rep3) %>% filter(Timepoint_hrs<30) %>% 
   filter(Timepoint_hrs!=12) %>% filter(Timepoint_hrs!=3) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = Timepoint_hrs, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_wrap(.~CodeName, ncol=1)+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=1, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=4) +
  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=1) +
  geom_text(data=statsbydayCFUsformat, aes(label="*", y=2e6, color=NA, shape=NA), size=9, color="#0fc08e",
            show.legend=FALSE)+
  scale_y_continuous(trans = 'log10', breaks=c(1e5, 1e6), # limits=c(1e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_x_continuous(breaks=c(0,6,12, 24))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", 
     #  title="All 3 experiments together",
       color=NULL, shape=NULL, lty=NULL)+
  theme(strip.text = element_markdown(size = 16, face="bold"), 
        panel.grid.minor.x = element_blank(),
        legend.text = element_text(size=20), legend.key.width = unit(2,"cm"),
         panel.background = element_rect(size=1, fill="white"))

survivalsublong<-dataallCFUs %>% pivot_longer(Rep1:Rep3) %>% filter(Timepoint_hrs==24) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = LoperamideTreatment, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_wrap(.~CodeName, ncol=1, scale="free_x")+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=0.2, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=4) +
  geom_text(data=statsbydayCFUsformat, aes(label="*", x=group2, y=2.5e6, color=NA, shape=NA), size=9, color="#0fc08e",
            show.legend=FALSE)+
  scale_y_continuous(trans = 'log10', breaks=c(1e5, 1e6), # limits=c(1e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
  labs(y="CFUs per mL water at 24 hr", x="Treatment", 
     #  title="All 3 experiments together",
       color=NULL, shape=NULL, lty=NULL)+
  theme(strip.text = element_markdown(size = 16, face="bold"), 
        panel.grid.minor.x = element_blank(),
        legend.text = element_text(size=20), legend.key.width = unit(2,"cm"),
        axis.line = element_line(size=0.5),axis.text.x = element_blank(),axis.ticks.x = element_blank(),
      panel.background = element_blank())
survivalsublong
#ggsave("WaterInVitro_LoperamideCFUs.png", bg="transparent", width = 12, height = 6)
```

### Timeline for each strain with mean until 24 hours - presentation style

```{r timelineplot24style, fig.height=5, fig.width=11}
dataallCFUs %>% pivot_longer(Rep1:Rep3) %>% filter(Timepoint_hrs<80) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = Timepoint_hrs, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment, group=LoperamideTreatment))+
  facet_wrap(.~Treatment, scales = "free_y", ncol=5)+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=1, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=1) +
  scale_y_continuous(trans = 'log10', limits=c(1e2,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  #scale_x_continuous(breaks=c(0,6,12,18,24))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", 
       color=NULL, shape=NULL, lty=NULL)+
   theme(text=element_text(family="Avenir"), #strip.text = element_markdown(size = 16), 
         axis.text = element_text(size=14), 
         axis.title = element_text(size=18), 
         legend.position = "bottom", legend.background = element_blank(),
         legend.text = element_text(size=18), legend.key.width = unit(1.4,"cm"),
         plot.background = element_blank(),
         strip.background = element_blank(),
         panel.background = element_rect(size=1, fill="white"))
#ggsave("WaterInVitro_LoperamideCFUs24avenir.png", bg="transparent", width = 16, height = 5)

```

### Timeline for each strain with mean until 24 hours

```{r timelineplot24, fig.height=5, fig.width=11}
dataallCFUs %>% pivot_longer(Rep1:Rep3) %>% filter(Timepoint_hrs<25) %>% 
  mutate(RepCFUs=(1000/VolPlated_ul)*DF*value) %>% 
  ggplot(aes(x = Timepoint_hrs, y=RepCFUs, 
             color=LoperamideTreatment, shape=LoperamideTreatment, group=LoperamideTreatment))+
  facet_grid(.~Treatment)+
  stat_summary(geom="errorbar", fun.data="mean_sd", width=1, show.legend = FALSE)+
  stat_summary(geom="point", fun="mean", size=3) +
  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=1) +
  scale_y_continuous(trans = 'log10', limits=c(5e4,1e7), 
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_x_continuous(breaks=c(0,6,12,18,24))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", 
       color=NULL, shape=NULL, lty=NULL)+
  theme(legend.key.width = unit(1.4,"cm"))
ggsave("WaterInVitro_LoperamideCFUs24.png", bg="transparent", width = 12, height = 5)
```


### Mean CFUs only

```{r waterplotall, fig.width=11, fig.height=5}
dataallCFUs %>% filter(Timepoint_hrs<80) %>% 
  ggplot(aes(x = Timepoint_hrs, y=CFUs_permL, 
             color=LoperamideTreatment, shape=LoperamideTreatment))+
  facet_grid(.~Treatment)+
  geom_point(size=3) +
  geom_smooth(aes(lty=LoperamideTreatment, fill=LoperamideTreatment), method="lm", na.rm=TRUE, alpha=0.2)+
#  stat_summary(aes(lty=LoperamideTreatment), geom="line", fun="mean", lwd=0.8) +
  scale_y_continuous(trans = 'log10', # expand=c(0,0),
                     labels = trans_format('log10', math_format(10^.x)))+
  scale_color_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
  scale_fill_manual(values=c('#000000', '#1c5580', '#0fc08e'))+
  scale_linetype_manual(values=c("solid","dotted","dashed"))+
  labs(y="CFUs per mL", x="Time since treatment (hours)", color=NULL, fill=NULL, shape=NULL, lty=NULL)+
  theme(legend.key.width = unit(1.33,"cm"))
```